---
title: "FamilyCode Figures"
output:
  html_document:
    df_print: paged
---

```{r}
#devtools::install_github('ShaoxunLiu/FCpackage')
library(FamilyCode)
```
# Load bHLH SELEX data
```{r}
library(FamilyCode)
workDir <- 'SELEX_bHLH/'
motifPoses <- c('P-3','P-2','P-1','P1','P2','P3')
```

1: Probound_runDir: Directory containing the ProBound results each in a folder.
2: bHLH_info: CSV table containing experiment information. 
3: bHLH_ali: Alignment file

```{r}
proBound_runDIR <- paste0(workDir, 'HTdata/')
bHLH_info <- read.csv(file = paste0(workDir, "bHLH_info.csv"), row.names = 1)
```

```{r}
proBound_files <- list.files(proBound_runDIR)
modelFile_Template <- paste0(proBound_runDIR, '$modelFile$/result/fit.models.consensus.json')
rows <- c("A", "C", "G", "T")
blank <- data.frame(P00 = c(1,1,1,1), P0 = c(1,1,1,1))

#getting expeirment identifier from bHLH_info, may vary according to different naming scheme
bHLH_index <- bHLH_info[,1:2]
```

This step can be ignored if no quality measurements are performed, or when each sample protein is only appearing once.

```{r}
bHLH_index$quality <- 0
for(i in 1:nrow(bHLH_index)){
  tryCatch({
    dir <- paste0(proBound_runDIR,bHLH_index$gene_symbol[i],'_',bHLH_index$study[i])
    suppressWarnings(quality <- readChar(paste0(dir,'//fitEval.txt'),nchars = 100))
    quality <- strsplit(quality, ', ')[[1]][1]
    quality <- strsplit(quality, ' = ')[[1]][2]
    bHLH_index$quality[i] <- quality
  }, error = function(e){})
}
```


Reads in all data of bHLH samples and compares the consensus sequence recognition score between the binding models output from ProBound. For bHLH, a logical value is given for symmetry. The motif with the highest consensus sequence recognition score is saved in the list of mono_motifs. The mono_motifs contains binding motifs in frequency format.

```{r}
Motif_Table <- loadMono_motifs(bHLH_index, modelFile_Template, rec_seq = 'CANNTG', pos_index = motifPoses, checkSymmetry = T, withTable = T)

all_mono_motifs <- Motif_Table$motifs
bHLH_model_info <- Motif_Table$table
```


Input Alignment file generated with Clustal Omega and select the aligned DNA-binding Domain of the family.

```{r}
bHLH_pbAlignment <- read.table(paste0(workDir, 'bHLH.sto'), quote="\"", fill = T)
name <- apply(bHLH_pbAlignment,1,function(x) strsplit(x[1],'/')[[1]][1])
colnames(bHLH_pbAlignment) <- c('name','alignment')
bHLH_pbAlignment$name <- name
```

For a different sample set, likely a different family, You can start from this point with properly structured list of mono_motifs and data frame of Alignment


Family code uses the bHLH_index data frame to load models. Filter rows according to recorded parameters and reload mono_motifs. Here, the data frame: bHLH_model_info has contains parameters and the first two rows are in the same format as the bHLH_index data frame.

```{r}
#filter for symmetry
bHLH_sym <- bHLH_model_info[bHLH_model_info$symmetry == 1,]
#filter for best model of each protein
bHLH_names <- unique(bHLH_sym$gene_symbol)
bHLH_index <- data.frame(NULL)
for(i in 1:length(bHLH_names)){
  qualities <- bHLH_sym[bHLH_sym$gene_symbol == bHLH_names[i],]
  add <- qualities[qualities$quality == max(qualities$quality),]
  bHLH_index <- rbind.data.frame(bHLH_index, add[1,])
}
#filter for quality > 0.15
bHLH_index <- bHLH_index[bHLH_index$quality > 0.15,]


#reload mono_motifs
bHLH_motifs <- loadMono_motifs(bHLH_index, modelFile_Template, rec_seq = 'CANNTG', pos_index = motifPoses, checkSymmetry = F, withTable = F)
#uniform variable name
CV_motif <- bHLH_motifs
CV_alignment <- matchAliMotif(CV_motif, bHLH_pbAlignment)
```

# Figure 1
```{r Figure 1}
library(readr)
library(R.utils)

mkdirs('logos')
for(i in 1:nrow(bHLH_index)){
  JSON_Lines <- readLines(paste0("HTdata/", bHLH_index$gene_symbol[i], '_', bHLH_index$study[i],'/result/fit.models.consensus.json'))
  mono <- JSON2Matrix(JSON_Lines, mode = bHLH_index$motif_model[i])
  plot <- mononucleotide_logo(mono, axes = F)
  png(paste0('logos/',bHLH_index$gene_symbol[i],'.png'))
  plot(plot)
  dev.off()
}
```

# Figure 2

```{r 2A}
library(plotly)
AAPosition <- 13
motifPosition <- 'P-1'
AAbp_Matrix <- AAbpCombination(CV_motif, CV_alignment, AAPosition, motifPosition)
plot <- plot_tetrahedron(AAbp_Matrix, color = T, vertex = F, size = 10)
scene = list(camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)))
plot%>%layout(scene=scene)
```

```{r 2B}
library(plotly)
AAPosition <- 5
motifPosition <- 'P-1'
AAbp_Matrix <- AAbpCombination(CV_motif, CV_alignment, AAPosition, motifPosition)
plot <- plot_tetrahedron(AAbp_Matrix, color = T, vertex = F, size = 10)
scene = list(camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)))
plot%>%layout(scene=scene)
```

```{r 2C}
library(plotly)
AAPosition <- 13
motifPosition <- 'P-3'
AAbp_Matrix <- AAbpCombination(CV_motif, CV_alignment, AAPosition, motifPosition)
plot <- plot_tetrahedron(AAbp_Matrix, color = T, vertex = F, size = 10)
scene = list(camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)))
plot%>%layout(scene=scene)
```

```{r 2D, S7A}
#genreate p-val tablle
pvalTable <- getPvalTable(CV_motif, Alignment = CV_alignment, pos_index = motifPoses)
#apply -log(10)
pvalTable <- -log(pvalTable,10)
#set NAs to 0
pvalTable[is.na(pvalTable)] <- 0
#plot heatmap
gplots::heatmap.2(as.matrix(as.data.frame(lapply(pvalTable, as.numeric))),dendrogram='none',
                  Rowv=FALSE, Colv=FALSE,trace='none',col = rev(heat.colors(12)), key.title = '-logP-val')
```


# Figure S1
```{r S1}
colorCode <- pvalTable$`P-1`
colorCode[is.na(colorCode)] <- median(colorCode[!is.na(colorCode)])
max <- max(colorCode)
colorList <- colorCode/max*255
colorList <- c(colorList[1:29],0,0,0,0,colorList[30:53])

complexSample <- 'bHLH_processed.pse'
addColorTemplate <- 'color 0x$colorCode$, polymer.protein and resi $resiNum$'
script <- paste('load ', complexSample, sep = '')
for(i in 1:length(colorList)){
  colorCode <- rgb(colorList[i], 0, (255-colorList[i]), maxColorValue = 255)
  colorCode <- gsub('#','',colorCode)
  addScript <- gsub('\\$colorCode\\$',colorCode,addColorTemplate)
  addScript <- gsub('\\$resiNum\\$', i, addScript)
  script <- c(script, addScript)
}
run.pymol(pymol.dir = 'E:/pymol/pymol_app/PyMOLWin.exe',script = script)

```

# Figure S2
```{r S2B}
AAPosition <- 20
motifPosition <- 'P-1'
AAbp_Matrix <- AAbpCombination(CV_motif, CV_alignment, AAPosition, motifPosition)
plot <- plot_tetrahedron(AAbp_Matrix, color = T, vertex = F, size = 10)
scene = list(camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)))
plot%>%layout(scene=scene)

#AAs <- unique(colnames(AAbp_Matrix))
#plot(1)
#legend(x = "topleft", legend = AAs, col = AAcolor(AAs), pch = 19, cex = 1, bty = 'n')
```

```{r S2D}
posMatrix <- gene2pos(CV_motif, pos = 'P-1')
P22 <- substr(CV_alignment$alignment,22, 22)
P50 <- substr(CV_alignment$alignment,50, 50)

table(AAclassify(P22), AAclassify(P50))
Encode22 <- attr(AAclassify(P22), 'encode')
Encode50 <- attr(AAclassify(P50), 'encode')
Interactions <- AAinteraction(Encode22, Encode50)
table(Interactions, apply(posMatrix, 2, function(x) rownames(posMatrix)[which.max(x)]))

paste0(Interactions, '=>',apply(posMatrix, 2, function(x) rownames(posMatrix)[which.max(x)]))
paste0(AAclassify(P22), '+',AAclassify(P50),'=',Interactions)
paste0(P22, '+',P50,'=',Interactions,  '=>',apply(posMatrix, 2, function(x) rownames(posMatrix)[which.max(x)]))

colnames(posMatrix) <- Interactions
colnames(posMatrix)[colnames(posMatrix) == 'h-bond'] <- 'Q'
colnames(posMatrix)[colnames(posMatrix) == 'electro-static'] <- 'E'
colnames(posMatrix)[colnames(posMatrix) == 'hydrophobic'] <- 'L'
colnames(posMatrix)[colnames(posMatrix) == 'repel'] <- 'W'

plot <- plot_tetrahedron(posMatrix, color = T, vertex = F, size = 10)
scene = list(camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)))
plot%>%layout(scene=scene)
#plot(1)

#legend(x = "topleft", legend = c('H-bond','Electro-static', 'Hydrophobic', 'Repel'), col = AAcolor(c('Q', 'E', 'L', 'W')), pch = 19, cex = 1, bty = 'n')
```

# Figure 3
```{r 3A}
library(plotly)
AAPosition <- 13
motifPosition <- 'P-1'
AAbp_Matrix <- AAbpCombination(CV_motif, CV_alignment, AAPosition, motifPosition)
Rmatirx <- AAbp_Matrix[,colnames(AAbp_Matrix) == 'R']
RTM <- matrix2tetrahedron(Rmatirx)
meanPointR <- apply((RTM), 2, function(x) mean(x))
Rmean <- tetrahedron2matrix(meanPointR)
svd <- FamilyCode::matrixSVD(AAbp_Matrix)
Allmean <- tetrahedron2matrix(svd$tetra_mean)
AAbp_Matrix <- AAbp_Matrix[,colnames(AAbp_Matrix) == 'R']
AAbp_Matrix <- cbind.data.frame(AAbp_Matrix, X = Allmean)
connect <- cbind.data.frame(Rmean = meanPointR, Allmean = svd$tetra_mean)
scene = list(camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)))

plotTest <- plot_tetrahedron(AAbp_Matrix, color = T, size = 10)
plotTest%>%
  add_trace(connect,x = as.numeric(connect[1,1:2]),
            y=as.numeric(connect[2,1:2]),z= as.numeric(connect[3,1:2]),color = I('black'),
            type = 'scatter3d',
            mode = 'lines',
            line = list(width = 5),
            opacity = 1,
            showlegend = F
  )%>%
  add_trace(
    type= "cone",
    x= connect[1,1], y= connect[2,1], z= connect[3,1],
    u= connect[1,1] - connect[1,2], v= connect[2,1] - connect[2,2], w= connect[3,1] - connect[3,2],
    sizeref= 0.5,
    colorscale = list(list(0, as.character('black')),
                      list(1, as.character('black'))),
    anchor = 'tip',
    showscale= F
  )%>%layout(scene=scene)

rownames(Rmean) <- c('A','C','G','T')
mononucleotide_logo(Rmean)

```

```{r 3B}
library(plotly)
AAPosition <- 13
motifPosition <- 'P-1'
AAbp_Matrix <- AAbpCombination(CV_motif, CV_alignment, AAPosition, motifPosition)
Rmatirx <- AAbp_Matrix[,colnames(AAbp_Matrix) == 'V']
RTM <- matrix2tetrahedron(Rmatirx)
meanPointR <- apply((RTM), 2, function(x) mean(x))
Rmean <- tetrahedron2matrix(meanPointR)
svd <- FamilyCode::matrixSVD(AAbp_Matrix)
Allmean <- tetrahedron2matrix(svd$tetra_mean)
AAbp_Matrix <- AAbp_Matrix[,colnames(AAbp_Matrix) == 'V']
AAbp_Matrix <- cbind.data.frame(AAbp_Matrix, X = Allmean)
connect <- cbind.data.frame(Rmean = meanPointR, Allmean = svd$tetra_mean)
scene = list(camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)))

plotTest <- plot_tetrahedron(AAbp_Matrix, color = T, size = 10)
plotTest%>%
  add_trace(connect,x = as.numeric(connect[1,1:2]),
            y=as.numeric(connect[2,1:2]),z= as.numeric(connect[3,1:2]),color = I('black'),
            type = 'scatter3d',
            mode = 'lines',
            line = list(width = 10),
            opacity = 1,
            showlegend = F
  )%>%
  add_trace(
    type= "cone",
    x= connect[1,1], y= connect[2,1], z= connect[3,1],
    u= connect[1,1] - connect[1,2], v= connect[2,1] - connect[2,2], w= connect[3,1] - connect[3,2],
    sizeref= 0.5,
    colorscale = list(list(0, as.character('black')),
                      list(1, as.character('black'))),
    anchor = 'tip',
    showscale= F
  )%>%layout(scene=scene)

rownames(Rmean) <- c('A','C','G','T')
mononucleotide_logo(log(Rmean))

```

```{r 3C}
AAPosition <- 13
motifPosition <- 'P-1'
AAbp_Matrix <- AAbpCombination(CV_motif, CV_alignment, AAPosition, motifPosition)

svd <- FamilyCode::matrixSVD(AAbp_Matrix)
vCoord <- svd$v
meanPointAll <- svd$tetra_mean
connect <- matrix(nrow = 4, ncol = 3, data = 0)
connect[4,] <- meanPointAll
connect[1,] <- connect[4,] + vCoord[,1]
connect[2,] <- connect[4,] + vCoord[,2]
connect[3,] <- connect[4,] + vCoord[,3]


tetrahedron <-  matrix(data = c(1,1,1,1,-1,-1,-1,1,-1,-1,-1,1,1,1,1,-1,1,-1,1,-1,-1,-1,-1,1), nrow = 8, ncol = 3, byrow = TRUE)
tetrahedron <- as.data.frame(tetrahedron)

plotTest <- plot_tetrahedron(AAbp_Matrix, color = T, size = 10)


plotTest <- plotTest%>%
  add_trace(tetrahedron, x = connect[c(1,4),1],
            y=connect[c(1,4),2],z=connect[c(1,4),3],color = I('grey'),
            type = 'scatter3d',
            mode = 'lines',
            line = list(width = 5),
            opacity = 0.5,
            showlegend = F
  )%>%
  add_trace(
    type= "cone",
    x= connect[1,1], y= connect[1,2], z= connect[1,3],
    u= connect[1,1] - connect[4,1], v= connect[1,2] - connect[4,2], w= connect[1,3] - connect[4,3],
    sizeref= 0.2,
    colorscale = list(list(0, as.character('grey')),
                      list(1, as.character('grey'))),
    anchor = 'tip',
    showscale= F
  )%>%
  add_trace(tetrahedron, x = connect[c(2,4),1],
            y=connect[c(2,4),2],z=connect[c(2,4),3],color = I('grey'),
            type = 'scatter3d',
            mode = 'lines',
            line = list(width = 5),
            opacity = 0.5,
            showlegend = F
  )%>%
  add_trace(
    type= "cone",
    x= connect[2,1], y= connect[2,2], z= connect[2,3],
    u= connect[2,1] - connect[4,1], v= connect[2,2] - connect[4,2], w= connect[2,3] - connect[4,3],
    sizeref= 0.2,
    colorscale = list(list(0, as.character('grey')),
                      list(1, as.character('grey'))),
    anchor = 'tip',
    showscale= F
  )%>%
  add_trace(tetrahedron, x = connect[c(3,4),1],
            y=connect[c(3,4),2],z=connect[c(3,4),3],color = I('grey'),
            type = 'scatter3d',
            mode = 'lines',
            line = list(width = 5),
            opacity = 0.5,
            showlegend = F
  )%>%
  add_trace(
    type= "cone",
    x= connect[3,1], y= connect[3,2], z= connect[3,3],
    u= connect[3,1] - connect[4,1], v= connect[3,2] - connect[4,2], w= connect[3,3] - connect[4,3],
    sizeref= 0.2,
    colorscale = list(list(0, as.character('grey')),
                      list(1, as.character('grey'))),
    anchor = 'tip',
    showscale= F
  )%>%
  add_text(x = connect[c(1:3),1], y = connect[c(1:3),2], z = connect[c(1:3),3], text = c('PC1', 'PC2','PC3'),
           showlegend = F)

  axx <- list(
    title = "",
    zeroline = TRUE,
    showline = TRUE,
    showticklabels = TRUE,
    showgrid = TRUE
  )
scene = list(camera = list(eye = list(x = 0, y = 0, z = -2)))

plotTest%>%layout(scene=scene)
```

#Figure S3
```{r 3D, S3B}
#make predictions
FCpredTrueList <- list()
pos <- 'P-1'

predTrue <- SVDregression.Iterative.CV(CV_motif,CV_alignment, pos = pos,Ftest_pVal = 0.001, useSimilarAA = T)
#delete unsuccessful predictions
predTrue <- predTrue[!is.na(predTrue$pred),]
FCpredTrueList[[pos]] <- predTrue

predTrue <- FCpredTrueList[[pos]]
#delete unsuccessful predictions
#predTrue <- predTrue[!is.na(predTrue$pred),]
attr(predTrue, 'confidence') <- attr(predTrue, 'confidence')[attr(predTrue, 'confidence') <= 1]
plotPredTrue(predTrue, main = paste0('FamilyCode at ',pos))

```
# Figure S4
```{r S4A}
#plot confidance
plot(attr(predTrue, 'confidence'),groupedR2(predTrue$pred, predTrue$true,member = 4, mean = F), ylab = 'Predict-True R^2', xlab = 'Confidence score', main = paste0('Confidence at ', pos))
legend('bottomright', paste0('R^2 = ', round(groupedR2(groupedR2(predTrue$pred, predTrue$true,member = 4, mean = F),
                                                   attr(predTrue, 'confidence'), ),4)), bty = 'n')
dt <- data.frame(conf = attr(predTrue, 'confidence'), R2 = groupedR2(predTrue$pred, predTrue$true,member = 4, mean = F))
lm <- lm(R2~conf, data= dt)
abline(lm, col= 'red')
#plot frequency

```

```{r S3A}
pos <- 'P-1'

BaseLines.ddG <- getBaseLineAccuracy(all_mono_motifs, pos = pos, randomSample = F)
colnames(BaseLines.ddG) <- c('pred','true')
BaseLines.PFM <- predTrue.ddG2frequency(BaseLines.ddG, PFM = T)
plotPredTrue(BaseLines.ddG, xlab = 'Experiment 1 -ΔΔG/RT', ylab = 'Experiment 2 -ΔΔG/RT', main = paste0('Pair-wise experiment comparison at ',pos))

```

```{r S3C}
pos <- 'P-1'

predTrue <- closestSeqPred(CV_motif,CV_alignment, pos = pos)
#delete unsuccessful predictions
predTrue <- predTrue[!is.na(predTrue$pred),]
plotPredTrue(predTrue, main = paste0('Closest-paralog prediction at ',pos))
```

```{r S3D}
pos <- 'P-1'

mod_alignment <- CV_alignment
mod_alignment$alignment <- paste0(substr(mod_alignment$alignment,1,34), '-','-',substr(mod_alignment$alignment,35,53))
predTrue <- SRpred(CV_motif,mod_alignment, pos = pos, weightfile = 'cisBP_bHLH/F082_1.97d.json')


plotPredTrue(predTrue, main = paste0('Similarity regression at ',pos))

```

#Figure 4
```{r}
groups <- c('HWT', 'H5', 'H13','HD', 'AWT', 'A5', 'A13', 'AD')
SELEXFile_Template <- 'SELEXdata/$modelFile$/fit.models.consensus.json'
SELEX_info <- data.frame(gene_symbol = groups, study = 'pb')
SELEX_motifs <- loadMono_motifs(SELEX_info, SELEXFile_Template, rec_seq = 'CANNTG', pos_index = c('P-3','P-2','P-1','P1','P2','P3'), checkSymmetry = T, withTable = F, useMode = c(1,2,1,1,1,1,1,1))
```

```{r 4B}
result <- readLines(paste0("HTdata/HES2_Yin2017/result/fit.models.consensus.json"))
mono <- JSON2Matrix(result, mode = 1)
mononucleotide_logo(frequency2ddG(mono))
result <- readLines(paste0("HTdata/ASCL2_Yin2017/result/fit.models.consensus.json"))
mono <- JSON2Matrix(result, mode = 1)
mononucleotide_logo(frequency2ddG(mono))
```

```{r 4C}
result <- readLines(paste0("SELEXdata/HWT_pb/fit.models.consensus.json"))
mono <- JSON2Matrix(result, mode = 1)
mononucleotide_logo(frequency2ddG(mono))
result <- readLines(paste0("SELEXdata/AWT_pb/fit.models.consensus.json"))
mono <- JSON2Matrix(result, mode = 1)
mononucleotide_logo(frequency2ddG(mono))
test <- data.frame(pred = as.numeric(frequency2ddG(JSON2Matrix(readLines(paste0("SELEXdata/HWT_pb/fit.models.consensus.json"))))[,4:6]), true = as.numeric(frequency2ddG(JSON2Matrix(readLines(paste0("HTdata/HES2_Yin2017/result/fit.models.consensus.json"))))[,4:6]))
plotPredTrue(test,  xlab = 'HT-SELEX -ΔΔG/RT', ylab = 'Experimental -ΔΔG/RT', main = paste0('HES2 replicate'))
test <- data.frame(pred = as.numeric(frequency2ddG(JSON2Matrix(readLines(paste0("SELEXdata/AWT_pb/fit.models.consensus.json"))))[,4:6]), true = as.numeric(frequency2ddG(JSON2Matrix(readLines(paste0("HTdata/ASCL2_Yin2017/result/fit.models.consensus.json"))))[,4:6]))
plotPredTrue(test,  xlab = 'HT-SELEX -ΔΔG/RT', ylab = 'Experimental -ΔΔG/RT', main = paste0('ASCL2 replicate'))
```


```{r 4E}
library(readxl)
library(ggplot2)
library(ggsignif)
bandIntensity <- read_excel("bandIntensity.xlsx")

proportions <- bandIntensity$...6[c(1:6,13:18)]
proportions <- c(proportions[1:3], exp(1.788), 8.09,
                 proportions[4:6],exp(1.956), 28.41,
                 proportions[7:9], exp(-3.948),0.023,
                 proportions[10:12], exp(-2.15), 0.135)
names <- c(rep('WT',5),rep('R5K',5),rep('V13R',5),rep('Double',5))
names <- as.factor(names)
cols <- c(rep('#6cab42',1),rep('#87eff2',1),rep('#f9d85b',1),rep('#c4560c',1))
dt <- data.frame(name = names, prop = proportions)
dt$prop <- log(dt$prop)
dt$name <- factor(dt$name, levels = c('WT', 'R5K', 'V13R', 'Double'))


p <- ggplot(dt, aes(x = name, y = prop, fill = name, color = name)) +
  geom_dotplot(binaxis = 'y', stackdir = 'center', dotsize = 1) +
  scale_fill_manual(values = cols, name = "Name") +
  scale_color_manual(values = cols, name = "Name") +
  labs(
    x = "",
    y = "-ΔΔG/RT (GC->CG)",
    size =15
  ) +
  theme(legend.position = "none",
    axis.text.x = element_text(face = 'bold',color=cols, size = 15),
    panel.background = element_blank(),     # Remove panel background
    panel.grid.major = element_blank(),     # Remove major gridlines
    panel.grid.minor = element_blank(),     # Remove minor gridlines
    plot.background = element_blank(),
    panel.border = element_rect(colour = "black", fill=NA, size=0.5)# Remove plot background
  ) +
  geom_signif(comparisons = list(c("R5K", "WT"),
                                 c("V13R", "Double"),
                                 c("R5K", "V13R"),
                                 c("WT", "V13R"),
                                 c("WT", "Double"),
                                 c("R5K", "Double")
                                 ),
              map_signif_level=TRUE, color = 'black',  step_increase = 0.1)
plot(p)
```


```{r 4FG}
svd <- matrixSVD(gene2pos(CV_motif, pos = 'P-1'))
svdModelpreTrain <- trainSVD.Iterative(svd, CV_alignment, Ftest_pVal = 0.001)
keyPos <- svdModelpreTrain$keyPos
keyPos <- lapply(keyPos, function(x) c(x,5))
#svdModel <- trainSVD(svd, CV_alignment, no.keyPos = unlist(lapply(keyPos,length)), keyPos = keyPos)
svdModel <- trainSVD(svd, CV_alignment, no.keyPos = c(2,1,1), keyPos = list(c(13,5), c(13), c(13)))
#Predict binding motifs for test set
test_alignment <- data.frame(name = paste0(groups, '_pb')[-c(1,5)], alignment =c(
                             'RKSLRPLLEKRRRARINQSLSQLKGLLPENSNCSKLEKADVLEMTVRFLQEL-',
                             'RKSLKPLLEKRRVARINQSLSQLKGLLPENSNCSKLEKADVLEMTVRFLQEL-',
                             'RKSLRPLLEKRRVARINQSLSQLKGLLPENSNCSKLEKADVLEMTVRFLQEL-',
                             '--ARKNERERNRVKLVNLGFQALRQHVPHGGASKKLSKVETLRSAVEYIRALQ',
                             '--ARRNERERNRRKLVNLGFQALRQHVPHGGASKKLSKVETLRSAVEYIRALQ',
                             '--ARKNERERNRRKLVNLGFQALRQHVPHGGASKKLSKVETLRSAVEYIRALQ')
                             )
test_motifs <- SELEX_motifs[-c(1,5)]
pred_motifs <- predict(svdModel, test_alignment[], zero = 0.01, useSimilarAA = T)
#Comparing between true and predicted testing set motifs 

predT <- matrix2tetrahedron(pred_motifs)

true <- lapply(test_motifs, function(x) x$matrix[,c('P-1','P-1')])
trueT <- lapply(true, matrix2tetrahedron)
trueT <- matrix(nrow = 6, ncol = 6, data = unlist(trueT), byrow = T)
trueT <- trueT[,c(1,3,5)]
wt <- lapply(SELEX_motifs[c(1,5)], function(x) x$matrix[,c('P-1','P-1')])
wtT <- lapply(wt, matrix2tetrahedron)
wtT <- matrix(nrow = 2, ncol = 6, data = unlist(wtT), byrow = T)
wtT <- wtT[,c(1,3,5)]
pred <- predT[1:6,] - wtT[c(1,1,1,2,2,2),]
true <- trueT[1:6,] - wtT[c(1,1,1,2,2,2),]
pred <- tetrahedron2matrix(pred)
true <- tetrahedron2matrix(true)
pred[pred < 0.01] <- 0.01
true[true < 0.01] <- 0.01

pred <- frequency2ddG(pred)
true <- frequency2ddG(true)
predTrueHes2 <- data.frame(pred = as.numeric(unlist(pred[,1:3])), true = as.numeric(unlist(true[,1:3])))
plotPredTrue(predTrueHes2,  xlab = 'Experimental -ΔΔΔG/RT', ylab = 'Predicted -ΔΔΔG/RT', main = paste0('FamilyCode on HES2 mutants'))
predTrueAscl2 <- data.frame(pred = as.numeric(unlist(pred[,4:6])), true = as.numeric(unlist(true[,4:6])))
plotPredTrue(predTrueAscl2,  xlab = 'Experimental -ΔΔΔG/RT', ylab = 'Predicted -ΔΔΔG/RT', main = paste0('FamilyCode on ASCL2 mutants'))
```

#Figure S5
```{r S5}
library(plotly)
AAbp_Matrix1 <- data.frame(Y = c(0.2,0.2,0.2,1))
AAbp_Matrix2 <- data.frame(H = c(0.1,0.1,1,0.1))
scene = list(camera = list(eye = list(x = 1.25, y = 1.25, z = 1.25)))

#300/500
rownames(AAbp_Matrix1) <- DNA()
rownames(AAbp_Matrix2) <- DNA()
mononucleotide_logo(frequency2ddG(AAbp_Matrix1))
mononucleotide_logo(frequency2ddG(AAbp_Matrix2))

mononucleotide_logo(get.dddG(AAbp_Matrix2, AAbp_Matrix1))

Tetra1 <- matrix2tetrahedron(AAbp_Matrix1)
Tetra2 <- matrix2tetrahedron(AAbp_Matrix2)
connect <- t(rbind.data.frame(Tetra2,Tetra1))


plotTest <- plot_tetrahedron(cbind(AAbp_Matrix1,AAbp_Matrix2), color = T, size = 20)
plotTest%>%
  add_trace(connect,x = as.numeric(connect[1,1:2]),
            y=as.numeric(connect[2,1:2]),z= as.numeric(connect[3,1:2]),color = I('black'),
            type = 'scatter3d',
            mode = 'lines',
            line = list(width = 5),
            opacity = 1,
            showlegend = F
  )%>%
  add_trace(
    type= "cone",
    x= connect[1,1], y= connect[2,1], z= connect[3,1],
    u= connect[1,1] - connect[1,2], v= connect[2,1] - connect[2,2], w= connect[3,1] - connect[3,2],
    sizeref= 0.1,
    colorscale = list(list(0, as.character('black')),
                      list(1, as.character('black'))),
    anchor = 'tip',
    showscale= F
  )%>%layout(scene=scene)


```

# Load bHLH PBM data
```{r}
library(FamilyCode)
library(R.utils)
workDir <- 'PBM_bHLH/'
#download from cisBP 'bulk download' tab, inducing Z-score and TF_Information
TF_Information <- read.delim(paste0(workDir, "TF_Information.txt"))
Zscores <- read.delim(paste0(workDir,"Zscores.txt"))
#filter for PBM data with Z-scores
TFDirect <- TF_Information[TF_Information$TF_Status == 'D',]
rawPBM <- substr(colnames(Zscores)[-1], 1, 11)
#deal with duplicated experiment
substr(rawPBM[duplicated(rawPBM)],11,11) <- '1'
colnames(Zscores)[-1] <- rawPBM
TFDirect.duplicated <- TFDirect
substr(TFDirect.duplicated$Motif_ID,11,11) <- '1'
TFDirect <- rbind.data.frame(TFDirect, TFDirect.duplicated)
#get cisBP entries with PBM Z-scores
TFDirect <- TFDirect[match(intersect(TFDirect$Motif_ID, rawPBM),TFDirect$Motif_ID),]


#filter out demasi mutation sets
TFdemasi <- TFDirect[TFDirect$MSource_Identifier == 'DeMasi2011',]
```

```{r}
#uniprot can only handle certain number of searches at a time, so seperate all data into two parts
gene_ids <- unique(TFDirect[,c('TF_Name', 'TF_Species')])
searchUnitTemplate <- '(gene:"GENE" AND "SPECIES")'
searchQ <- gsub('GENE', gene_ids$TF_Name[1], searchUnitTemplate)
searchQ <- gsub('SPECIES', gene_ids$TF_Species[1], searchQ)
for(i in 1:nrow(gene_ids)){#nrow(gene_ids)
  addQ <- gsub('GENE', gene_ids$TF_Name[i], searchUnitTemplate)
  addQ <- gsub('SPECIES', gene_ids$TF_Species[i], addQ)
  searchQ <- paste0(searchQ, ' OR ', addQ)
}
```

```{r}
geneInfoUniprot <- read.delim(paste0(workDir,"geneInfoUniprot_all.tsv"))
geneInfoUniprot$Gene.Names <- paste0(geneInfoUniprot$Gene.Names, ' ')
```

```{r}

geneInfo <- data.frame()
for(i in 1:nrow(gene_ids)){
  add <- geneInfoUniprot[grep(paste0(gene_ids$TF_Name[i],' '), geneInfoUniprot$Gene.Names),]
  if(nrow(add) != 0){
    add <- add[nchar(add$Sequence) == max(nchar(add$Sequence)),][1,]
    addSeq <- add$Sequence
  }else{
    addSeq <- NA 
  }
  Motif <- unique(TFDirect[TFDirect$TF_Name == gene_ids$TF_Name[i],]$Motif_ID)
  addLine <- data.frame(name = gene_ids$TF_Name[i], seq = addSeq, motif = Motif)
  geneInfo <- rbind.data.frame(geneInfo, addLine)
}

geneInfo <- geneInfo[!is.na(geneInfo$seq),]
#write sequences as fasta to
#writeFasta(geneInfo, paste0(workDir,'TF.fa'))
```

```{r}
ali <- read.table(paste0(workDir,"TFhmm.sto"), quote="\"", fill = T)
ali <- ali[-nrow(ali),]
ali$V1 <- apply(ali, 1, function(x) strsplit(x[1],'/')[[1]][1])
ali$V1 <- apply(ali, 1, function(x) strsplit(x[1],'\\|')[[1]][2])
colnames(ali) <- c('name','alignment')
keep <- which(substr(ali$alignment, 9,9) == 'E')
cisBP_alignment <- ali[keep,]
geneInfo <- geneInfo[keep,]
#Check if cisBP_alignment and geneInfo matches
sum(cisBP_alignment$name != geneInfo$name) == 0
```

```{r}
geneInfo$name[match(TFdemasi$Motif_ID, geneInfo$motif)] <- paste0(TFdemasi$DBID.1, 'DeMasi')
```

```{r}
cisBP_motifs <- list()
options(warn=0)
for(i in 1:nrow(geneInfo)){
  motif <- geneInfo$motif[i]
  motif <- suppressWarnings(read.table(paste0(workDir, 'pwms/', motif, '.txt'), header = F))
  motif <- ddG2frequency(motif)
  motif[motif < 0.001] <- 0.001#add regularization
  rownames(motif) <- DNA()
  add <- list()
  add$gene_symbol <- geneInfo$name[i]
  add$study <- i
  add$mode <- 1
  add$matrix <- as.matrix(motif)
  cisBP_motifs[[length(cisBP_motifs)+1]] <- add
  add <- list()
  #add reverse compliment motif
  add$gene_symbol <- geneInfo$name[i]
  add$study <- i
  add$mode <- -1
  if(ncol(motif) == 0){
    rev <- motif
  }else{
    rev <- as.matrix(motif)[nrow(motif):1, ncol(motif):1]
  }
  rownames(rev) <- DNA()
  add$matrix <- rev
  cisBP_motifs[[length(cisBP_motifs)+1]] <- add
}
#check if length of cisBP_motif is equal to 2 times geneInfo
length(cisBP_motifs) == nrow(geneInfo)*2
DeMasiID <- which(unlist(lapply(cisBP_motifs, function(x) length(grep('DeMasi', x$gene_symbol)) > 0)))
DeMasi_motifs <- cisBP_motifs[DeMasiID]
cisBP_motifs <- cisBP_motifs[-DeMasiID]

```

```{r}
#align to CANNTG motif
bHLHseed <- matrix(nrow = 4,ncol = 6,data = c(0,1,0,0,
                                              1,0,0,0,
                                              0,0,0,0,
                                              0,0,0,0,
                                              0,0,0,1,
                                              0,0,1,0
))
rownames(bHLHseed) <- DNA()
#Scoring would result in some motif shorter than the seed, those motifs will be discarded. 
motifScores <- scoreMotifList(cisBP_motifs, bHLHseed, weight = c(1,1,0,0,1,1))
```

The motif (forward or reverse compliment) with the lowest motif matching score of each entry will be kept. For the same protein, entries are arranged from lower score to higher score
```{r}
idstds <- unique(motifScores[,1:2])
all_motifs <- data.frame(NULL)
for(i in 1:nrow(idstds)){
  fDT <- motifScores[motifScores$gene_symbol == idstds[i,1],]
  fDT <- fDT[fDT$study == idstds[i,2],]
  all_motifs <- rbind.data.frame(all_motifs, fDT[which.min(fDT$score),])
}
motifs_arranged <- data.frame(NULL)
motifScore <- c()
for(i in 1:length(unique(all_motifs$gene_symbol))){
  add <- all_motifs[all_motifs$gene_symbol == unique(all_motifs$gene_symbol)[i],]
  add <- dplyr::arrange(add, score)
  motifScore <- c(motifScore,add$score[1])
  motifs_arranged <- rbind.data.frame(motifs_arranged, add)
}
#Plot distribution of lowest motif matching score
#plot.ecdf(as.numeric(motifScore))
```

```{r}
#Filter out motifs that do not match well with the seed
motifTable_baseline <- motifs_arranged[motifs_arranged$score <= 1,]
#Number of unique HD protiens kept:
length(unique(motifTable_baseline$gene_symbol))
```

# Figure S6
```{r S6A}
motifPoses <- c('P-3','P-2','P-1','P1','P2','P3')
cisBP_motifs_baseline <- filterMotifList(motifTable_baseline, cisBP_motifs, 6, motifPoses)
cisBP_alignment <- cisBP_alignment[match(motifTable_baseline$gene_symbol, cisBP_alignment$name),]
#check if motif List has same length as alignment
length(cisBP_motifs_baseline) == nrow(cisBP_alignment)

pos <- 'P-1'
ScreenPoses <- c(pos,pos)
BaseLines.ddG <- getBaseLineAccuracy(cisBP_motifs_baseline, pos = ScreenPoses, randomSample = F)
colnames(BaseLines.ddG) <- c('pred','true')
plotPredTrue(BaseLines.ddG, xlab = 'Experiment 1 -ΔΔG/RT', ylab = 'Experiment 2 -ΔΔG/RT', main = paste0('Pair-wise experiment comparison at ',pos))

```

# Figure S7
```{r S7B}
del <- which(duplicated(cisBP_alignment$alignment))
CV_alignment <- cisBP_alignment[-del,]
CV_motif <- cisBP_motifs_baseline[-del]
#genreate p-val tablle
pvalTable <- getPvalTable(CV_motif, Alignment = CV_alignment, pos_index = motifPoses)
#apply -log(10)
pvalTable <- -log(pvalTable,10)
#set NAs to 0
pvalTable[is.na(pvalTable)] <- 0
#plot heatmap
#pvalTable[pvalTable > -log(0.001/53/3,10)] <- 10
gplots::heatmap.2(as.matrix(as.data.frame(lapply(pvalTable, as.numeric))),dendrogram='none',
                  Rowv=FALSE, Colv=FALSE,trace='none',col = rev(heat.colors(12)), key.title = '-logP-val')
```

```{r S6B}
pos <- 'P-1'
predTrue <- SVDregression.Iterative.CV(CV_motif,CV_alignment, pos = pos,Ftest_pVal = 0.001, useSimilarAA = T)
predTrue <- predTrue[!is.na(predTrue$pred),]
plotPredTrue(predTrue, main = paste0('FamilyCode at ', pos))
```

```{r S4B}
plot(attr(predTrue, 'confidence'),groupedR2(predTrue$pred, predTrue$true,member = 4, mean = F), ylab = 'Predict-True R^2', xlab = 'Confidence score', main = paste0('Confidence at ', pos))
legend('bottomright', paste0('R^2 = ', round(groupedR2(groupedR2(predTrue$pred, predTrue$true,member = 4, mean = F),
                                                   attr(predTrue, 'confidence'), ),4)), bty = 'n')
dt <- data.frame(conf = attr(predTrue, 'confidence'), R2 = groupedR2(predTrue$pred, predTrue$true,member = 4, mean = F))
lm <- lm(R2~conf, data= dt)
abline(lm, col= 'red')
```

```{r S6C}
predTrue <- closestSeqPred(CV_motif,CV_alignment, pos = pos)
#delete unsuccessful predictions
predTrue <- predTrue[!is.na(predTrue$pred),]

plotPredTrue(predTrue, main = paste0('Closest-paralog prediction at ',pos))
  
```

```{r S6D}

mod_alignment <- CV_alignment
mod_alignment$alignment <- paste0(substr(mod_alignment$alignment,1,34), '-','-',substr(mod_alignment$alignment,35,53))
predTrue <- SRpred(CV_motif,mod_alignment, pos = pos, weightfile = 'cisBP_bHLH/F082_1.97d.json')
#delete unsuccessful predictions
predTrue <- predTrue[!is.na(predTrue$pred),]


plotPredTrue(predTrue, main = paste0('Similarity regression at ',pos))

```

# Figure S8
```{r}
DeMasiInfo <- data.frame(name = TFdemasi$DBID.1, seq = 'RRKAATMRERRRRRKVNEAFEVVKQRTC-PNPNQRLPKVEILRSAIDYINNLE', motif = TFdemasi$Motif_ID)
substr(DeMasiInfo$seq,13,13) <- c('R','T','V','L')


#Scoring would result in some motif shorter than the seed, those motifs will be discarded. 
DeMasimotifScores <- scoreMotifList(DeMasi_motifs, bHLHseed, weight = c(1,1,0,0,1,1))

idstds <- unique(DeMasimotifScores[,1:2])
Demasi_motifsDT <- data.frame(NULL)
for(i in 1:nrow(idstds)){
  fDT <- DeMasimotifScores[DeMasimotifScores$gene_symbol == idstds[i,1],]
  fDT <- fDT[fDT$study == idstds[i,2],]
  Demasi_motifsDT <- rbind.data.frame(Demasi_motifsDT, fDT[which.min(fDT$score),])
}

DeMasi_motifs <- filterMotifList(Demasi_motifsDT, DeMasi_motifs, 6, motifPoses)
```

```{r S8B}
for(i in 1:4){
  plot <- mononucleotide_logo(frequency2ddG(DeMasi_motifs[[i]]$matrix))
  plot <- plot + ggplot2::ggtitle(paste0(DeMasiInfo$name[i],'_true'))
  plot(plot)
}
```

create FamilyCode model
```{r}
demasi_alignment <- rbind.data.frame(CV_alignment, as.character(DeMasiInfo[4,c(1:2)]))
demasi_motif <- CV_motif
demasi_motif[[84]] <- DeMasi_motifs[[4]]
svdModeldemasi <- makeSVDModel(CV_motif, CV_alignment, Positions = c('P-3','P-2','P-1','P1','P2','P3'))

```

predict ΔΔΔG using tetrahedron distance

```{r S8C}
svd <- matrixSVD(gene2pos(CV_motif, pos = 'P-1'))
#svdModel <- trainSVD(svd, CV_alignment, no.keyPos = unlist(lapply(keyPos,length)), keyPos = keyPos)
svdModel <- trainSVD(svd, CV_alignment, no.keyPos = c(1,1,1), keyPos = list(c(13), c(13), c(13)))
#Predict binding motifs for test set
test_alignment <- data.frame(name = DeMasiInfo$name, alignment = DeMasiInfo$seq)
pred_motifs <- predict(svdModel, test_alignment[], zero = 0.001, useSimilarAA = T)
#Comparing between true and predicted testing set motifs 

predT <- matrix2tetrahedron(pred_motifs)

true <- lapply(DeMasi_motifs, function(x) x$matrix[,c('P-1','P-1')])
trueT <- lapply(true, matrix2tetrahedron)
trueT <- matrix(nrow = 8, ncol = 6, data = unlist(trueT), byrow = T)
trueT <- trueT[,c(1,3,5)]
pred <- predT[1:4,] - trueT[c(4,4,4,4),]
true <- trueT[1:4,] - trueT[c(4,4,4,4),]
pred <- tetrahedron2matrix(pred)
true <- tetrahedron2matrix(true)
pred[pred < 0.01] <- 0.01#min(pred[pred > 0])
true[true < 0.01] <- 0.01#min(true[true > 0])

pred <- frequency2ddG(pred)
true <- frequency2ddG(true)
predTrue <- data.frame(pred = as.numeric(unlist(pred[,1:3])), true = as.numeric(unlist(true[,1:3])))
plotPredTrue(predTrue, main = 'FamilyCode prediction for HLH-1 mutants', xlab = 'DeMasi et. al., -ΔΔΔG/RT', ylab = 'Predicted -ΔΔΔG/RT')

```
```{r S8D}
HLH1WT <- read.delim("E:/Desktop/FamilyCode/deepPBS/HLH1/HLH1WT.tsv", header=FALSE)
HLH1L13V <- read.delim("E:/Desktop/FamilyCode/deepPBS/HLH1/HLH1L13V.tsv", header=FALSE)
HLH1L13R <- read.delim("E:/Desktop/FamilyCode/deepPBS/HLH1/HLH1L13R.tsv", header=FALSE)
HLH1L13T <- read.delim("E:/Desktop/FamilyCode/deepPBS/HLH1/HLH1L13T.tsv", header=FALSE)

PBSmodels <- list(HLH1L13R, HLH1L13T, HLH1L13V, HLH1WT, HLH1L13R, HLH1L13T, HLH1L13V, HLH1WT)
pred_motifs <- data.frame(NULL)
for(i in 1:length(PBSmodels)){
  PBSmotif <- t(PBSmodels[[i]])
  rownames(PBSmotif) <- DNA()
  PBSmotif <- apply(PBSmotif, 2, function(x) x/max(x))
  pred_motifs <- rbind.data.frame(pred_motifs, PBSmotif[,15])
}

pred_motifs <- t(pred_motifs)
rownames(pred_motifs) <- DNA()
#Comparing between true and predicted testing set motifs 

predT <- matrix2tetrahedron(pred_motifs)

true <- lapply(DeMasi_motifs, function(x) x$matrix[,c('P-1','P-1')])
trueT <- lapply(true, matrix2tetrahedron)
trueT <- matrix(nrow = 8, ncol = 6, data = unlist(trueT), byrow = T)
trueT <- trueT[,c(1,3,5)]
pred <- predT[1:4,] - trueT[c(4,4,4,4),]
true <- trueT[1:4,] - trueT[c(4,4,4,4),]
pred <- tetrahedron2matrix(pred)
true <- tetrahedron2matrix(true)
pred[pred < 0.01] <- 0.01#min(pred[pred > 0])
true[true < 0.01] <- 0.01#min(true[true > 0])

pred <- frequency2ddG(pred)
true <- frequency2ddG(true)
predTrue <- data.frame(pred = as.numeric(unlist(pred[,1:3])), true = as.numeric(unlist(true[,1:3])))
plotPredTrue(predTrue, main = 'DeepPBS prediction for HLH-1 mutants', xlab = 'DeMasi et. al., -ΔΔΔG/RT', ylab = 'Predicted -ΔΔΔG/RT')

```

# Load HD SELEX data
```{r}
library(FamilyCode)
workDir <- 'SELEX_HD/'
motifPoses <- c('N1','N2','T3','D4','A5','Y6','N7','N8')
#loadMotifs
proBound_DIR <- paste0(workDir, 'consensus_fits/')
template <- paste0(workDir, 'consensus_fits/$modelFile$.consensus.json')
gene_symbol_factor_ids <- read.csv(paste0(workDir, 'gene_symbol_factor_ids.csv'), row.names=1)
files <- list.files(proBound_DIR, pattern = 'consensus')
files <- unlist(lapply(strsplit(files, '\\.'), function(x) x[[1]][1]))
gene_symbol <- unlist(lapply(strsplit(files, '_'), function(x) x[[1]][1]))
study <- unlist(lapply(strsplit(files, '_'), function(x) x[[2]][1]))
info <- data.frame(gene_symbol, study)
info <- info[as.numeric(info$gene_symbol) < 10000,]

HD_all_motifs <- loadFromIndex(info, template)
motifPositions <- motifPoses
HDseed <- matrix(nrow = 4,ncol = 8,data = c(0,0,0,0,
                                            0,0,0,0,
                                            0,0,0,1,
                                            1,0,1,1,
                                            1,0,0,0,
                                            0,1,0,1,
                                            0,0,0,0,
                                            0,0,0,0))
rownames(HDseed) <- DNA()

HD_motifScores <- scoreMotifList(HD_all_motifs, HDseed, weight = c(0,0,1,1,1,1,0,0))

HDids<- unique(HD_motifScores$gene_symbol)
HDfiltered_table <- data.frame(NULL)
for(i in 1:length(HDids)){
  fDT <- HD_motifScores[HD_motifScores$gene_symbol == HDids[i],]
  HDfiltered_table <- rbind.data.frame(HDfiltered_table, fDT[which.min(fDT$score),])
}
plot.ecdf(as.numeric(HDfiltered_table$score))
HDfiltered_table <- HDfiltered_table[HDfiltered_table$score < 1,]

#prepare alignment
HD_ali <- read.table(paste0(workDir,"HD_hmm.sto"), quote="\"", fill = T)
HD_ali$V1 <- apply(HD_ali, 1, function(x) strsplit(x[1],'/')[[1]][1])
colnames(HD_ali) <- c('name','alignment')

#change names
geneName <- apply(HDfiltered_table, 1, function(x) gene_symbol_factor_ids$gene_symbol[gene_symbol_factor_ids$factor_ids == x[1]])
HDfiltered_table$gene_symbol <- geneName
HD_motifs <- filterMotifList(HDfiltered_table, HD_all_motifs, 8,motifPositions)


#quality filter
HD_alignment <- matchAliMotif(HD_motifs, HD_ali)
del <- which(substr(HD_alignment$alignment, 50,50) != 'N')
HD_alignment <- HD_alignment[-del,]
HD_motifs <- HD_motifs[-del]


del <- which(duplicated(HD_alignment$name))
HD_alignment_SELEX <- HD_alignment[-del,]
HD_motif_SELEX <- HD_motifs[-del]
```

```{r}
#get base line data set
HDidstds <- unique(HD_motifScores[,1:2])
HDfiltered_table_baseline <- data.frame(NULL)
for(i in 1:nrow(HDidstds)){
  fDT <- HD_motifScores[HD_motifScores$gene_symbol == HDidstds[i,1],]
  fDT <- fDT[fDT$study == HDidstds[i,2],]
  HDfiltered_table_baseline <- rbind.data.frame(HDfiltered_table_baseline, fDT[which.min(fDT$score),])
}
plot.ecdf(as.numeric(HDfiltered_table_baseline$score))

HDfiltered_table_baseline <- HDfiltered_table_baseline[HDfiltered_table_baseline$score < 1,]
HD_motifs_baseline <- filterMotifList(HDfiltered_table_baseline, HD_all_motifs, 8, motifPositions)

```

```{r S4C}
#genreate p-val tablle
pvalTable <- getPvalTable(HD_motif_SELEX, Alignment = HD_alignment_SELEX, pos_index = motifPositions)
#apply -log(10)
pvalTable <- -log(pvalTable,10)
#set NAs to 0
pvalTable[is.na(pvalTable)] <- 0
#plot heatmap
gplots::heatmap.2(as.matrix(as.data.frame(lapply(pvalTable, as.numeric))),dendrogram='none',
                  Rowv=FALSE, Colv=FALSE,trace='none',col = rev(heat.colors(12)), key.title = '-logP-val')

```

# Load HD PBM data
```{r}
library(FamilyCode)
library(R.utils)
workDir <- 'PBM_HD/'
#download from cisBP 'bulk download' tab, inducing Z-score and TF_Information
TF_Information <- read.delim(paste0(workDir, "TF_Information.txt"))
Zscores <- read.delim(paste0(workDir,"Zscores.txt"))
#filter for PBM data with Z-scores
TFDirect <- TF_Information[TF_Information$TF_Status == 'D',]
rawPBM <- substr(colnames(Zscores)[-1], 1, 11)
#deal with duplicated experiment
substr(rawPBM[duplicated(rawPBM)],11,11) <- '1'
colnames(Zscores)[-1] <- rawPBM
TFDirect.duplicated <- TFDirect
substr(TFDirect.duplicated$Motif_ID,11,11) <- '1'
TFDirect <- rbind.data.frame(TFDirect, TFDirect.duplicated)
#get cisBP entries with PBM Z-scores
TFDirect <- TFDirect[match(intersect(TFDirect$Motif_ID, rawPBM),TFDirect$Motif_ID),]

geneInfoUniprot <- read.delim(paste0(workDir,"geneInfoUniprot_all.tsv"))
geneInfoUniprot$Gene.Names <- paste0(geneInfoUniprot$Gene.Names, ' ')

gene_ids <- unique(TFDirect[,c('TF_Name', 'TF_Species')])
geneInfo <- data.frame()
for(i in 1:nrow(gene_ids)){
  add <- geneInfoUniprot[grep(paste0(gene_ids$TF_Name[i],' '), geneInfoUniprot$Gene.Names),]
  if(nrow(add) != 0){
    add <- add[nchar(add$Sequence) == max(nchar(add$Sequence)),][1,]
    addSeq <- add$Sequence
  }else{
    addSeq <- NA 
  }
  Motif <- unique(TFDirect[TFDirect$TF_Name == gene_ids$TF_Name[i],]$Motif_ID)
  addLine <- data.frame(name = gene_ids$TF_Name[i], seq = addSeq, motif = Motif)
  geneInfo <- rbind.data.frame(geneInfo, addLine)
}

geneInfo <- geneInfo[!is.na(geneInfo$seq),]


ali <- read.table(paste0(workDir,"TFhmm.sto"), quote="\"", fill = T)
ali <- ali[-nrow(ali),]
ali$V1 <- apply(ali, 1, function(x) strsplit(x[1],'/')[[1]][1])
ali$V1 <- apply(ali, 1, function(x) strsplit(x[1],'\\|')[[1]][2])
colnames(ali) <- c('name','alignment')
keep <- which(substr(ali$alignment, 50,50) == 'N')
cisBP_alignment <- ali[keep,]
geneInfo <- geneInfo[keep,]
#Check if cisBP_alignment and geneInfo matches
sum(cisBP_alignment$name != geneInfo$name) == 0
```
```{r}
cisBP_motifs <- list()
options(warn=0)
for(i in 1:nrow(geneInfo)){
  motif <- geneInfo$motif[i]
  motif <- suppressWarnings(read.table(paste0(workDir, 'pwms/', motif, '.txt'), header = F))
  motif <- ddG2frequency(motif)
  motif[motif < 0.001] <- 0.001#add psuedo count
  rownames(motif) <- DNA()
  add <- list()
  add$gene_symbol <- geneInfo$name[i]
  add$study <- i
  add$mode <- 1
  add$matrix <- as.matrix(motif)
  cisBP_motifs[[length(cisBP_motifs)+1]] <- add
  add <- list()
  #add reverse compliment motif
  add$gene_symbol <- geneInfo$name[i]
  add$study <- i
  add$mode <- -1
  if(ncol(motif) == 0){
    rev <- motif
  }else{
    rev <- as.matrix(motif)[nrow(motif):1, ncol(motif):1]
  }
  rownames(rev) <- DNA()
  add$matrix <- rev
  cisBP_motifs[[length(cisBP_motifs)+1]] <- add
}
#check if length of cisBP_motif is equal to 2 times geneInfo
length(cisBP_motifs) == nrow(geneInfo)*2

#align to NNTDAYNN motif
HDseed <- matrix(nrow = 4,ncol = 8,data = c(0,0,0,0,
                                            0,0,0,0,
                                            0,0,0,1,
                                            1,0,1,1,
                                            1,0,0,0,
                                            0,1,0,1,
                                            0,0,0,0,
                                            0,0,0,0))
rownames(HDseed) <- DNA()

motifScores <- scoreMotifList(cisBP_motifs, HDseed, weight = c(0,0,1,1,1,1,0,0))

idstds <- unique(motifScores[,1:2])
all_motifs <- data.frame(NULL)
for(i in 1:nrow(idstds)){
  fDT <- motifScores[motifScores$gene_symbol == idstds[i,1],]
  fDT <- fDT[fDT$study == idstds[i,2],]
  all_motifs <- rbind.data.frame(all_motifs, fDT[which.min(fDT$score),])
}
motifs_arranged <- data.frame(NULL)
motifScore <- c()
for(i in 1:length(unique(all_motifs$gene_symbol))){
  add <- all_motifs[all_motifs$gene_symbol == unique(all_motifs$gene_symbol)[i],]
  add <- dplyr::arrange(add, score)
  motifScore <- c(motifScore,add$score[1])
  motifs_arranged <- rbind.data.frame(motifs_arranged, add)
}
#Plot distribution of lowest motif matching score
#plot.ecdf(as.numeric(motifScore))

#Filter out motifs that do not match well with the seed
motifTable_baseline <- motifs_arranged[motifs_arranged$score < 1,]
#Number of unique HD protiens kept:
length(unique(motifTable_baseline$gene_symbol))

motifPoses <- c('N1','N2','T3','D4','A5','Y6','N7','N8')
cisBP_motifs_baseline <- filterMotifList(motifTable_baseline, cisBP_motifs, 8, motifPoses)
cisBP_alignment <- cisBP_alignment[match(motifTable_baseline$gene_symbol, cisBP_alignment$name),]
#check if motif List has same length as alignment
length(cisBP_motifs_baseline) == nrow(cisBP_alignment)

del <- which(duplicated(cisBP_alignment$alignment))
HD_alignment_PBM <- cisBP_alignment[-del,]
HD_motif_PBM <- cisBP_motifs_baseline[-del]
```

```{r S4D}
#genreate p-val tablle
pvalTable <- getPvalTable(HD_motif_PBM, Alignment = HD_alignment_PBM, pos_index = motifPositions)
#apply -log(10)
pvalTable <- -log(pvalTable,10)
#set NAs to 0
pvalTable[is.na(pvalTable)] <- 0
#plot heatmap
gplots::heatmap.2(as.matrix(as.data.frame(lapply(pvalTable, as.numeric))),dendrogram='none',
                  Rowv=FALSE, Colv=FALSE,trace='none',col = rev(heat.colors(12)), key.title = '-logP-val')

```


# Concatnate SELEX and PBM
```{r}
intersect <- intersect(HD_alignment_PBM$alignment, HD_alignment_SELEX$alignment)
PBM_int <- match(intersect, HD_alignment_PBM$alignment)
SELEX_int <- match(intersect, HD_alignment_SELEX$alignment)
all_motif <- c(HD_motif_SELEX, HD_motif_PBM)
CV_alignment <- rbind.data.frame(HD_alignment_SELEX, HD_alignment_PBM[-PBM_int,])
CV_motif <- c(HD_motif_SELEX, HD_motif_PBM[-PBM_int])
nrow(CV_alignment) == length(CV_motif)
saveRDS(CV_alignment, file = 'HD_SELEX_PBM_CV_alignment.rsd')
saveRDS(CV_motif, file = 'HD_SELEX_PBM_CV_motif.rsd')
CV_alignment <- readRDS('HD_SELEX_PBM_CV_alignment.rsd')
CV_motif <- readRDS('HD_SELEX_PBM_CV_motif.rsd')
```

# Figure 5
```{r 5A}
#genreate p-val tablle
pvalTable <- getPvalTable(CV_motif, Alignment = CV_alignment, pos_index = motifPoses)
#apply -log(10)
pvalTable <- -log(pvalTable,10)
#set NAs to 0
pvalTable[is.na(pvalTable)] <- 0
#plot heatmap
gplots::heatmap.2(as.matrix(as.data.frame(lapply(pvalTable, as.numeric))),dendrogram='none',
                  Rowv=FALSE, Colv=FALSE,trace='none',col = rev(heat.colors(12)), key.title = '-logP-val')
```

```{r 5B}

AAPosition <- 5
motifPosition <- 'D4'
AAbp_Matrix <- AAbpCombination(CV_motif, CV_alignment, AAPosition, motifPosition)
plot_tetrahedron(AAbp_Matrix, color = T, vertex = F, size = 10)
```


# Figure S9
```{r S9}
library(R.utils)
mkdirs('HDlogos')
for(i in 1:length(CV_motif)){
  mono <- CV_motif[[i]]$matrix
  plot <- mononucleotide_logo(frequency2ddG(mono), axes = F, labels = F)
  png(paste0('HDlogos/',CV_alignment$name[i],'_',substr(CV_alignment$alignment[i], 5,15), '-',substr(CV_alignment$alignment[i], 44,54),'.png'),
       width = 600, height = 200,)
  plot(plot)
  dev.off()
}
```

# Load Kock et al. data
```{r}
suppressMessages(library(readxl))
suppressMessages(library(FamilyCode))
suppressMessages(library(dplyr))
options(warn = -1)
workDir <- 'KockHDPBM/'
KockMatrix <- readRDS(paste0(workDir,'IntensityMatrix.rds'))
col <- gsub('NKX2-','NKX2.',colnames(KockMatrix[,3:ncol(KockMatrix)]))
col <- gsub('NKX3-','NKX3.',col)
CV_alignment <- readRDS('HD_SELEX_PBM_CV_alignment.rsd')
CV_motif <- readRDS('HD_SELEX_PBM_CV_motif.rsd')
motifPoses <- c('N1','N2','T3','D4','A5','Y6','N7','N8')

options(warn = -1)
ali <- read.table(paste0(workDir,"add_hmm.sto"), quote="\"", fill = T)
ali <- ali[-nrow(ali),]
ali$V1 <- apply(ali, 1, function(x) strsplit(x[1],'/')[[1]][1])
ali$V1 <- apply(ali, 1, function(x) strsplit(x[1],'\\|')[[1]][2])
colnames(ali) <- c('name','alignment')
ali$name <- c('HOXB6', 'NKX2-6', 'NKX2-4')
CV_alignment <- rbind.data.frame(CV_alignment, ali)

CV_motif[[415]] <- list(name = 'HOXB6', matrix = CV_motif[[414]]$matrix)
CV_motif[[416]] <- list(name = 'NKX2-6', matrix = CV_motif[[414]]$matrix)
CV_motif[[417]] <- list(name = 'NKX2-4', matrix = CV_motif[[414]]$matrix)

SampleInfo <- read_excel("KockHDPBM/41467_2024_47396_MOESM5_ESM.xls")
ExpIDs <- unique(unlist(lapply(strsplit(col,'_'), function(x) x[1])))
mutationInfo <- data.frame(NULL)
for(i in 1:length(ExpIDs)){
  id <- ExpIDs[i]
  info <- strsplit(id,'-')[[1]]
  if(info[2] == 'REF'){
    addLine <- data.frame(TF = info[1], mutation = 'WT', position = NA, mutateFrom = 'WT')
  }else{
    id <- gsub('\\.','-',id)
    pos <- SampleInfo[SampleInfo$allele == id,]$canonical
    addLine <- data.frame(TF = info[1], mutation = substr(info[2],nchar(info[2]),nchar(info[2])), position = pos, mutateFrom = substr(info[2],1,1))
  }
  mutationInfo <- rbind.data.frame(mutationInfo, addLine)
}

#validate with CV_Alignment
valid <- c()
mutationInfo$TF <- gsub('\\.','-',mutationInfo$TF)
for(i in 1:nrow(mutationInfo)){
  seq <- CV_alignment[CV_alignment$name == mutationInfo$TF[i],]$alignment
  if(mutationInfo$mutation[i] == 'WT' ){
    valid <- c(valid, 'WT')
  }else if(length(seq) == 0){
    valid <- c(valid, 'notInAlignment')
  }else{
    match <- substr(seq[1], mutationInfo$position[i]-1,mutationInfo$position[i]-1) == mutationInfo$mutateFrom[i]
    valid <- c(valid, match)
  }
}
mutationInfo$valid <- valid

alignment <- c()
for(i in 1:nrow(mutationInfo)){
  seq <- CV_alignment[CV_alignment$name == mutationInfo$TF[i],]$alignment
  if(length(seq) == 0){
    alignment <- c(alignment, NA)
  }else if(mutationInfo$mutation[i] == 'WT' ){
    alignment <- c(alignment, seq[1])
  }else{
    #match <- substr(seq[1], mutationInfo$position[i]-1,mutationInfo$position[i]-1) == mutationInfo$mutateFrom[i]
    #if(match){
    substr(seq[1], mutationInfo$position[i]-1,  mutationInfo$position[i]-1) <-  mutationInfo$mutation[i]
    alignment <- c(alignment, seq[1])
    #}else{
    #  alignment <- c(alignment, NA)
    #}
  }
}
mutationInfo$alignment <- alignment
mutationInfo$name <- gsub('\\.','-',ExpIDs)

PWMsID <- colnames(KockMatrix)[3:ncol(KockMatrix)]
Kock_motifs <- list()
options(warn=-1)
for(i in 1:length(PWMsID)){
  motif <- PWMsID[i]
  motif <- suppressWarnings(read.table(paste0(workDir, 'pwmsI/', motif, '.txt'), header = F))
  motif <- ddG2frequency(motif)
  motif[motif < 0.01] <- 0.01#add psuedo count
  rownames(motif) <- DNA()
  add <- list()
  add$gene_symbol <- PWMsID[i]
  add$study <- i
  add$mode <- 1
  add$matrix <- as.matrix(motif)
  colnames(add$matrix) <- motifPoses
  Kock_motifs[[length(Kock_motifs)+1]] <- add
}


psamI_pcc <- read.table(paste0(workDir,'psamI_pcc.txt'), quote="\"", comment.char="")
colnames(psamI_pcc) <- c('sample','R2')

key <- match(PWMsID,psamI_pcc$sample)
psamI_pcc <- psamI_pcc[key,]
sum(psamI_pcc$sample == PWMsID)

R2order <- order(-psamI_pcc$R2)

psamI_pcc$name = unlist(lapply(psamI_pcc$sample, function(x) strsplit(x,'_')[[1]][1]))
psamI_pcc <- psamI_pcc[R2order,]
Kock_motifs <- Kock_motifs[R2order]

bestMatch <- which(!duplicated(psamI_pcc$name))
Kock_motifs_bestRep <- Kock_motifs[bestMatch]
psamI_pcc_bestRep <- psamI_pcc[bestMatch,]
psamI_pcc_2 <- psamI_pcc[-bestMatch,]
Kock_motifs_2 <- Kock_motifs[-bestMatch]
bestMatch2 <- which(!duplicated(psamI_pcc_2$name))
Kock_motifs_2bestRep <- Kock_motifs_2[bestMatch2]
psamI_pcc_2bestRep <- psamI_pcc_2[bestMatch2,]


#Construct motif list and alignment table for Base Line (Motif similarity between the two best PBM experiments)

Kock_alignment <- data.frame(name = mutationInfo$name, alignment = mutationInfo$alignment)

Kock_motifs_set <- list()
Exp2TrueMotifs <- list()
for(i in 1:nrow(Kock_alignment)){
  bestID <- which(psamI_pcc_bestRep$name == Kock_alignment$name[i])
  add <- list(name = Kock_alignment$name[i], probeR2 = psamI_pcc_bestRep$R2[bestID], matrix = Kock_motifs_bestRep[[bestID]]$matrix)
  Kock_motifs_set[[length(Kock_motifs_set)+1]] <- add
  bestID <- which(psamI_pcc_2bestRep$name == Kock_alignment$name[i])
  if(length(bestID) ==0){
    Exp2TrueMotifs[[length(Exp2TrueMotifs)+1]] <- list(name = Kock_alignment$name[i], probeR2 = NULL, matrix = NULL)
    next
  }
  add <- list(name = Kock_alignment$name[i], probeR2 = psamI_pcc_2bestRep$R2[bestID], matrix = Kock_motifs_2bestRep[[bestID]]$matrix)
  Exp2TrueMotifs[[length(Exp2TrueMotifs)+1]] <- add
}

#check if motif List has same length as alignment
length(Kock_motifs_set) == nrow(Kock_alignment)
sum(unlist(lapply(Kock_motifs_set, function(x) x$name)) == Kock_alignment$name)

sum(mutationInfo$name == Kock_alignment$name)

CV_alignment_Kock <- CV_alignment[1:414,]
CV_motif_Kock <- CV_motif[1:414]

#Prepare true motifs
mutantID <- which(mutationInfo$mutate != 'WT')
WTID <- which(mutationInfo$mutate == 'WT')
trueMotifs <- Kock_motifs_set[mutantID]
trueStandard <- list()
for(i in mutantID){
  setA <- which(mutationInfo$TF == mutationInfo$TF[i])
  setB <- which(mutationInfo$mutation == 'WT')
  WTid <- intersect(setA, setB)
  add <- Kock_motifs_set[[WTid[1]]]
  trueStandard[[length(trueStandard)+1]] <- add
}
```

#Figure 6
```{r baseLine}
dddGTrue <- list()
dddGpred <- list()
R2ListbaseLine <- c()
for(i in 1:length(trueMotifs)){
  mutid <- mutantID[i]
  if(is.null(Exp2TrueMotifs[[mutid]]$matrix)){
    R2ListbaseLine <- c(R2ListbaseLine, 0)
    dddGTrue[[length(dddGTrue) + 1]] <- NULL
    dddGpred[[length(dddGpred) + 1]] <- NULL
    next
  }
  predDiffTetra <- matrix2tetrahedron(Exp2TrueMotifs[[mutid]]$matrix) - matrix2tetrahedron(trueStandard[[i]]$matrix)
  #predDiffTetra <- matrix2tetrahedron(Exp2TrueMotifs[[mutid]]$matrix) - matrix2tetrahedron(Exp2TrueMotifs[[max(WTID[WTID<mutid])]]$matrix)
  if(sum(predDiffTetra) == 0){
    R2ListbaseLine <- c(R2ListbaseLine, 0)
    dddGTrue[[length(dddGTrue) + 1]] <- NULL
    dddGpred[[length(dddGpred) + 1]] <- NULL
    next
  }
  trueDiffTetra <- matrix2tetrahedron(trueMotifs[[i]]$matrix) - matrix2tetrahedron(trueStandard[[i]]$matrix)
  #trueDiffTetra <- matrix2tetrahedron(trueMotifs[[i]]$matrix) - matrix2tetrahedron(Exp2TrueMotifs[[max(WTID[WTID<mutid])]]$matrix)
  predDiffMatrix <- tetrahedron2matrix(predDiffTetra)
  trueDiffMatrix <- tetrahedron2matrix(trueDiffTetra)
  predDiffMatrix[predDiffMatrix < 0.01] <- 0.01
  trueDiffMatrix[trueDiffMatrix < 0.01] <- 0.01
  rownames(trueDiffMatrix) <- DNA()
  rownames(predDiffMatrix) <- DNA()
  pred <- as.numeric(frequency2ddG(predDiffMatrix))
  true <- as.numeric(frequency2ddG(trueDiffMatrix))
  predTrue <- data.frame(true = true, pred = pred)
  dddGTrue[[length(dddGTrue) + 1]] <- trueDiffMatrix
  dddGpred[[length(dddGpred) + 1]] <- predDiffMatrix
  R2ListbaseLine <- c(R2ListbaseLine, groupedR2(predTrue$true, predTrue$pred, member = nrow(predTrue)))
}
```

```{r FamilyCode}
mutantID <- which(mutationInfo$mutate != 'WT')
WTID <- which(mutationInfo$mutate == 'WT')

standard_Alignment <- list()
for(i in 1:nrow(Kock_alignment)){
  WTid <- which(CV_alignment$name == mutationInfo$TF[i])
  add <- CV_alignment[WTid[1],]
  standard_Alignment <- rbind.data.frame(standard_Alignment, add)
}

#WTmodel <- makeSVDModel(CV_motif_Kock, CV_alignment_Kock, Positions = motifPoses, Ftest_pVal = 0.01)

mutationPredList <- list()
mutationPredConfidence <- c()
inTraining <- c()
for(i in mutantID){
  mutpos <- as.numeric(mutationInfo$position[i])-1
  svdModel <- makeSVDModel.singleMutation(CV_motif_Kock, CV_alignment_Kock, Positions = motifPoses, mutation = mutpos)
  inTraining <- c(sum(substr(Kock_alignment$alignment[i],mutpos,mutpos) == substr(CV_alignment_Kock$alignment,mutpos,mutpos)), inTraining)
  #predict mutant
  predPSAM <- data.frame(pred = c(0,0,0,0))
  rownames(predPSAM) <- DNA()
  for(pos in motifPoses){
    predColumn <- predict(svdModel[[pos]], Kock_alignment[i,], useSimilarAA = T)
    predPSAM <- cbind.data.frame(predPSAM, predColumn)
    predConfidence <- attr(predColumn, 'confidence')
  }
  predPSAM <- predPSAM[,-1]
  colnames(predPSAM) <- motifPoses
  #predict standard
  
  mutationPredList[[length(mutationPredList)+1]] <- list(name = Kock_alignment$name[i], matrix = predPSAM)
  mutationPredConfidence <- c(mutationPredConfidence, predConfidence)
}


predMotifs <- mutationPredList
sum(unlist(lapply(predMotifs, function(x) x$name)) == unlist(lapply(trueMotifs, function(x) x$name)))

dddGTrue <- list()
dddGpred <- list()
R2List <- c()
for(i in 1:length(trueMotifs)){
  predDiffTetra <- matrix2tetrahedron(predMotifs[[i]]$matrix) - matrix2tetrahedron(trueStandard[[i]]$matrix)
  if(sum(predDiffTetra) == 0){
    predDiffTetra[1] <- 0.01
    mod <- TRUE
  }else{
    mod <- FALSE
  }
  trueDiffTetra <- matrix2tetrahedron(trueMotifs[[i]]$matrix) - matrix2tetrahedron(trueStandard[[i]]$matrix)
  predDiffMatrix <- tetrahedron2matrix(predDiffTetra)
  trueDiffMatrix <- tetrahedron2matrix(trueDiffTetra)
  predDiffMatrix[predDiffMatrix < 0.01] <- 0.01
  trueDiffMatrix[trueDiffMatrix < 0.01] <- 0.01
  rownames(trueDiffMatrix) <- DNA()
  rownames(predDiffMatrix) <- DNA()
  if(mod){
    predDiffMatrix[1,] <- 1
  }
  pred <- as.numeric(frequency2ddG(predDiffMatrix))
  true <- as.numeric(frequency2ddG(trueDiffMatrix))
  predTrue <- data.frame(true = true, pred = pred)
  dddGTrue[[length(dddGTrue) + 1]] <- trueDiffMatrix
  dddGpred[[length(dddGpred) + 1]] <- predDiffMatrix
  #plotPredTrue(predTrue, main = paste0('FamilyCode rediction for ',unlist(lapply(trueMotifs, function(x) x$name))[i]), xlab = 'Experimental -ΔΔΔG/RT', ylab = 'Predicted -ΔΔΔG/RT')
  R2List <- c(R2List, groupedR2(predTrue$true, predTrue$pred, member = nrow(predTrue)))
}
trueMagnitudeAll <- c()
predMagnitudeAll <- c()
for(pos in motifPoses){
  pred <- unlist(lapply(dddGpred, function(x) frequency2ddG(x)[,pos]))
  true <- unlist(lapply(dddGTrue, function(x) frequency2ddG(x)[,pos]))
  predTrue <- data.frame(true = true, pred = pred)
  predTrue <- predTrue[!is.na(predTrue$pred),]
  trueMatrix <- matrix(nrow = 4, ncol = nrow(predTrue)/4, byrow = F, data = predTrue$true)
  trueMagnitude <- c()
  predMatrix <- matrix(nrow = 4, ncol = nrow(predTrue)/4, byrow = F, data = predTrue$pred)
  predMagnitude <- c()
  for(set in 1:ncol(trueMatrix)){
    trueMagnitude <- c(trueMagnitude, max(trueMatrix[,set]) - min(trueMatrix[,set]))
    predMagnitude <- c(predMagnitude, max(predMatrix[,set]) - min(predMatrix[,set]))
  }
  trueMagnitudeAll <- c(trueMagnitudeAll, trueMagnitude)
  predMagnitudeAll <- c(predMagnitudeAll, predMagnitude)
}
  
groupedpredMagnitudeAll <- matrix(nrow = 8, ncol = 92, data = predMagnitudeAll, byrow = T)
dddGRangePred <- apply(groupedpredMagnitudeAll, 2, function(x) max(x))
groupedtrueMagnitudeAll <- matrix(nrow = 8, ncol = 92, data = trueMagnitudeAll, byrow = T)
dddGRangeTrue <- apply(groupedtrueMagnitudeAll, 2, function(x) max(x))
match <- match(mutationInfo$name[mutantID], SampleInfo$allele)
SampleInfo$allele[match] == mutationInfo$name[mutantID]
matchSampleInfo <- SampleInfo[match,]
```

```{r 6A, S10C}
# Create a data frame for plotting
data <- data.frame(
  position = mutationInfo$position[mutantID]-1,
  R2List = R2List,
  trueMagnitude = dddGRangeTrue
)

# Define the color categories
data$color_category <- ifelse(
  ((matchSampleInfo$affinity_change == 'no change') + (matchSampleInfo$specificity_change == 'no')) == 2, "Mutation has no effect", "Specificity or affinity change")

# Define color mapping
color_mapping <- c("Specificity or affinity change" = "red",  "Mutation has no effect" = "cyan")

# First Lollipop Plot: R2List
ggplot(data, aes(x = factor(position), y = R2List, color = color_category)) +
  geom_segment(aes(xend = factor(position), y = 0, yend = R2List), 
               color = "gray", linetype = "dotted", size = 0.5) +  # Gray, dotted, thin sticks
  geom_point(size = 3) +
  scale_color_manual(values = color_mapping) +
  labs(title = "",
       x = "Position of mutation of HD alignment",
       y = "",
       color = "Annotation according to Kock et al.") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),  # Remove background grid
    axis.text.x = element_text(angle = 90, vjust = 0.5)
  )

```

```{r 6BC}
i = which(mutationInfo[mutantID,]$position == 29)
i = i[which.max(dddGRangeTrue[i])]

predDiffTetra <- matrix2tetrahedron(predMotifs[[i]]$matrix) - matrix2tetrahedron(trueStandard[[i]]$matrix)
if(sum(predDiffTetra) == 0){
  predDiffTetra[1] <- 0.01
  mod <- TRUE
}else{
  mod <- FALSE
}
trueDiffTetra <- matrix2tetrahedron(trueMotifs[[i]]$matrix) - matrix2tetrahedron(trueStandard[[i]]$matrix)
predDiffMatrix <- tetrahedron2matrix(predDiffTetra)
trueDiffMatrix <- tetrahedron2matrix(trueDiffTetra)
predDiffMatrix[predDiffMatrix < 0.01] <- 0.01
trueDiffMatrix[trueDiffMatrix < 0.01] <- 0.01
rownames(trueDiffMatrix) <- DNA()
rownames(predDiffMatrix) <- DNA()
if(mod){
  predDiffMatrix[1,] <- 1
}
pred <- as.numeric(frequency2ddG(predDiffMatrix))
true <- as.numeric(frequency2ddG(trueDiffMatrix))
predTrue <- data.frame(true = true, pred = pred)
plotPredTrue(predTrue, main = paste0('FamilyCode prediction for ',unlist(lapply(trueMotifs, function(x) x$name))[i]), xlab = 'Replicate1 -ΔΔΔG/RT', ylab = 'Predicted -ΔΔΔG/RT')
mononucleotide_logo(frequency2ddG(predDiffMatrix), label = F)
mononucleotide_logo(frequency2ddG(trueDiffMatrix), label = F)
#mononucleotide_logo(frequency2ddG(Kock_motifs_set[[mut]]$matrix))
#mononucleotide_logo(frequency2ddG(Kock_motifs_set[[max(WTID[WTID < mut])]]$matrix))

mutid <- mutantID[i]
predDiffTetra <- matrix2tetrahedron(Exp2TrueMotifs[[mutid]]$matrix) - matrix2tetrahedron(trueStandard[[i]]$matrix)
#predDiffTetra <- matrix2tetrahedron(Exp2TrueMotifs[[mutid]]$matrix) - matrix2tetrahedron(Exp2TrueMotifs[[max(WTID[WTID<mutid])]]$matrix)
trueDiffTetra <- matrix2tetrahedron(trueMotifs[[i]]$matrix) - matrix2tetrahedron(trueStandard[[i]]$matrix)
#trueDiffTetra <- matrix2tetrahedron(trueMotifs[[i]]$matrix) - matrix2tetrahedron(Exp2TrueMotifs[[max(WTID[WTID<mutid])]]$matrix)
predDiffMatrix <- tetrahedron2matrix(predDiffTetra)
trueDiffMatrix <- tetrahedron2matrix(trueDiffTetra)
predDiffMatrix[predDiffMatrix < 0.01] <- 0.01
trueDiffMatrix[trueDiffMatrix < 0.01] <- 0.01
rownames(trueDiffMatrix) <- DNA()
rownames(predDiffMatrix) <- DNA()
pred <- as.numeric(frequency2ddG(predDiffMatrix))
true <- as.numeric(frequency2ddG(trueDiffMatrix))
predTrue <- data.frame(true = true, pred = pred)
plotPredTrue(predTrue, main = paste0('Base Line for ',unlist(lapply(trueMotifs, function(x) x$name))[i]), xlab = 'Replicate1 -ΔΔΔG/RT', ylab = 'Replicate2 -ΔΔΔG/RT')
mononucleotide_logo(frequency2ddG(predDiffMatrix), label = F)
mononucleotide_logo(frequency2ddG(trueDiffMatrix), label = F)
```

```{r rCLAMPS}
rCLAMPS_predicted_pwms <- read.delim("E:/Desktop/FamilyCode/KockHDPBM/rCLAMPS_predicted_pwms.txt")
motifPoses <- c('T3','D4','A5','Y6','N7','N8')
rCLAMPSTFs <- mutationInfo$name[mutationInfo$mutation != 'WT']
predMotifs <- list()
for(i in 1:length(rCLAMPSTFs)){
  motifInfo <- rCLAMPS_predicted_pwms[rCLAMPS_predicted_pwms$prot == rCLAMPSTFs[i],]
  matrix <- matrix(ncol = 6, nrow = 4, data = motifInfo$prob, byrow = F)
  rownames(matrix) <- DNA()
  colnames(matrix) <- motifPoses
  predMotifs[[length(predMotifs)+1]] <- list(name = motifInfo$prot[1], matrix = matrix)
}



sum(unlist(lapply(predMotifs, function(x) x$name)) == unlist(lapply(trueMotifs, function(x) x$name)))



dddGTrue <- list()
dddGpred <- list()
R2List_rCLAMPS <- c()
for(i in 1:length(trueMotifs)){
  predDiffTetra <- matrix2tetrahedron(predMotifs[[i]]$matrix) - matrix2tetrahedron(trueStandard[[i]]$matrix[,motifPoses])
  if(sum(predDiffTetra) == 0){
    next
  }
  trueDiffTetra <- matrix2tetrahedron(trueMotifs[[i]]$matrix[,motifPoses]) - matrix2tetrahedron(trueStandard[[i]]$matrix[,motifPoses])
  predDiffMatrix <- tetrahedron2matrix(predDiffTetra)
  trueDiffMatrix <- tetrahedron2matrix(trueDiffTetra)
  predDiffMatrix[predDiffMatrix < 0] <- 0.01
  trueDiffMatrix[trueDiffMatrix < 0] <- 0.01
  rownames(trueDiffMatrix) <- DNA()
  rownames(predDiffMatrix) <- DNA()
  pred <- as.numeric(frequency2ddG(predDiffMatrix))
  true <- as.numeric(frequency2ddG(trueDiffMatrix))
  predTrue <- data.frame(true = true, pred = pred)
  dddGTrue[[length(dddGTrue) + 1]] <- trueDiffMatrix
  dddGpred[[length(dddGpred) + 1]] <- predDiffMatrix
  R2List_rCLAMPS <- c(R2List_rCLAMPS, groupedR2(predTrue$true, predTrue$pred, member = nrow(predTrue)))
}
```

```{r deepPBS}
motifPoses <- c('N1','N2','T3','D4','A5','Y6','N7','N8')

deepPBSdir <- 'deepPBS/HDKock/results/'
deepPBSmotifFiles <- list.files(deepPBSdir, pattern = '.tsv')

deepPBSTFs <- gsub('.tsv','',deepPBSmotifFiles)
deepPBSTFs <- gsub('_','-',deepPBSTFs)
deepPBSdMotifs <- list()
for(i in 1:length(deepPBSmotifFiles)){
  PBSmodels <- read.delim(paste0(deepPBSdir, deepPBSmotifFiles[i]), header=FALSE)
  PBSmotif <- t(PBSmodels)
  motif <- apply(PBSmotif, 2, function(x) x/max(x))
  rownames(motif) <- DNA()
  add <- list()
  add$gene_symbol <- deepPBSTFs[i]
  add$study <- i
  add$mode <- 1
  add$matrix <- as.matrix(motif)
  deepPBSdMotifs[[length(deepPBSdMotifs)+1]] <- add
  add <- list()
  #add reverse compliment motif
  add$gene_symbol <- deepPBSTFs[i]
  add$study <- i
  add$mode <- -1
  if(ncol(motif) == 0){
    rev <- motif
  }else{
    rev <- as.matrix(motif)[nrow(motif):1, ncol(motif):1]
  }
  rownames(rev) <- DNA()
  add$matrix <- rev
  deepPBSdMotifs[[length(deepPBSdMotifs)+1]] <- add
}

HDseed <- matrix(nrow = 4,ncol = 8,data = c(0,0,0,0,
                                            0,0,0,0,
                                            0,0,0,1,
                                            1,0,1,1,
                                            1,0,0,0,
                                            0,1,0,1,
                                            0,0,0,0,
                                            0,0,0,0))
rownames(HDseed) <- DNA()

motifScores <- scoreMotifList(deepPBSdMotifs, HDseed, weight = c(0,0,1,1,1,1,0,0))
idstds <- unique(motifScores[,1:2])
all_motifs <- data.frame(NULL)
for(i in 1:nrow(idstds)){
  fDT <- motifScores[motifScores$gene_symbol == idstds[i,1],]
  fDT <- fDT[fDT$study == idstds[i,2],]
  all_motifs <- rbind.data.frame(all_motifs, fDT[which.min(fDT$score),])
}

all_motifs$study <- unlist(lapply(strsplit(all_motifs$gene_symbol,'_'), function(x) x[2]))
all_motifs$gene_symbol <- unlist(lapply(strsplit(all_motifs$gene_symbol,'_'), function(x) x[1]))
motifs_arranged <- data.frame(NULL)
motifScore <- c()
for(i in 1:length(unique(all_motifs$gene_symbol))){
  add <- all_motifs[all_motifs$gene_symbol == unique(all_motifs$gene_symbol)[i],]
  add <- dplyr::arrange(add, score)
  motifScore <- c(motifScore,add$score[1])
  motifs_arranged <- rbind.data.frame(motifs_arranged, add)
}
#Plot distribution of lowest motif matching score
plot.ecdf(as.numeric(motifScore))

deepPBS_motifs_set <- filterMotifList(motifs_arranged, deepPBSdMotifs, 8, motifPoses)
predMotifs <- deepPBS_motifs_set

key <- match(tolower(unlist(lapply(trueMotifs, function(x) x$name))),unlist(lapply(predMotifs, function(x) x$name)))
predMotifs <- predMotifs[key]
sum(unlist(lapply(predMotifs, function(x) x$name)) == tolower(unlist(lapply(trueMotifs, function(x) x$name))))

#dddG prediction
dddGTrue <- list()
dddGpred <- list()
R2List_deepPBS <- c()
for(i in 1:length(trueMotifs)){
  predDiffTetra <- matrix2tetrahedron(predMotifs[[i]]$matrix) - matrix2tetrahedron(trueStandard[[i]]$matrix)
  if(sum(predDiffTetra) == 0){
    next
  }
  trueDiffTetra <- matrix2tetrahedron(trueMotifs[[i]]$matrix) - matrix2tetrahedron(trueStandard[[i]]$matrix)
  predDiffMatrix <- tetrahedron2matrix(predDiffTetra)
  trueDiffMatrix <- tetrahedron2matrix(trueDiffTetra)
  predDiffMatrix[predDiffMatrix < 0] <- 0.01
  trueDiffMatrix[trueDiffMatrix < 0] <- 0.01
  rownames(trueDiffMatrix) <- DNA()
  rownames(predDiffMatrix) <- DNA()
  pred <- as.numeric(frequency2ddG(predDiffMatrix))
  true <- as.numeric(frequency2ddG(trueDiffMatrix))
  predTrue <- data.frame(true = true, pred = pred)
  dddGTrue[[length(dddGTrue) + 1]] <- trueDiffMatrix
  dddGpred[[length(dddGpred) + 1]] <- predDiffMatrix
  R2List_deepPBS <- c(R2List_deepPBS, groupedR2(predTrue$true, predTrue$pred, member = nrow(predTrue)))
}
```

```{r 6D}
library(ggplot2)
library(ggpubr)

rCmutpos <- 1:92
data <- data.frame(
  R2 = c(R2ListbaseLine[rCmutpos], R2List[rCmutpos], R2List_rCLAMPS[rCmutpos], R2List_deepPBS[rCmutpos]),
  Group = factor(c(rep("Base Line", length(rCmutpos)), 
                   rep("Family Code", length(rCmutpos)), 
                   rep("rCLAMPS", length(rCmutpos)), 
                   rep("deepPBS", length(rCmutpos))))
)

data$Group <- factor(data$Group, levels = c('Base Line', 'Family Code', 'rCLAMPS', 'deepPBS'))

# Custom colors for the groups
custom_colors <- c('Base Line' = '#1f77b4', 
                   'Family Code' = '#ff7f0e', 
                   'rCLAMPS' = '#2ca02c', 
                   'deepPBS' = '#d61828')

# Create violin plot with jittered points colored by group
p <- ggplot(data, aes(x = Group, y = R2, fill = Group)) +
  geom_violin(trim = FALSE, alpha = 0.6) +  # Violin plot
  geom_jitter(aes(color = Group), shape = 16, size = 1.5, position = position_jitter(width = 0.02, height = 0), alpha = 1) + # Jitter points with matching colors
  labs(title = "ΔΔΔG/RT Prediction R^2 Comparison", y = "R^2s", x = "") +
  scale_fill_manual(values = custom_colors) +  # Apply custom fill colors
  scale_color_manual(values = custom_colors) + # Apply matching colors to jitter points
  theme_minimal() +
  ylim(-0.5, 1.7) +
  theme(
    legend.position = "bottom",        # Place legend at the bottom
    legend.title = element_blank(),    # Remove legend title
    panel.grid.major = element_blank(),# Remove major grid lines
    panel.grid.minor = element_blank(),
    panel.border = element_blank(),    # Remove minor grid lines
    panel.background = element_rect(fill = "white"), # White panel background
    plot.background = element_rect(color = "white"),
    plot.title = element_text(face = "bold", size = 14, hjust = 0.5)
  )

# Add significance bars with fixed vertical lines
p <- p + stat_compare_means(comparisons = list(
                                               c("Family Code", "rCLAMPS"),
                                               c("Family Code", "deepPBS"),
                                               c('Base Line', "Family Code")),
                            label = "p.signif",
                            method = "wilcox",
                            paired = TRUE,
                            tip.length = 0.08, # Shorter vertical lines
                            bracket.size = 0.7,
                            label.y= c(1.3,1.45,1.6)) +  # Thicker brackets for clarity
  theme(legend.position = "none")

# Print the plot
plot(p)
```

# Figure S10
```{r S10A}
# Create a data frame for plotting
data <- data.frame(
  position = mutationInfo$position[mutantID]-1,
  R2List = R2ListbaseLine,
  trueMagnitude = dddGRangeTrue
)

# Define the color categories
# Define the color categories
data$color_category <- ifelse(
  ((matchSampleInfo$affinity_change == 'no change') + (matchSampleInfo$specificity_change == 'no')) == 2, "Mutation has no effect", "Specificity or affinity change")

# Define color mapping
color_mapping <- c("Specificity or affinity change" = "red",  "Mutation has no effect" = "cyan")
# Second Lollipop Plot: trueMagnitude
ggplot(data, aes(x = factor(position), y = trueMagnitude, color = color_category)) +
  geom_segment(aes(xend = factor(position), y = 0, yend = trueMagnitude), 
               color = "gray", linetype = "dotted", size = 0.5) +  # Gray, dotted, thin sticks
  geom_point(size = 3, alpha = 1) +
  #geom_hline(yintercept = 1, linetype = "dashed", color = "gray", size = 0.7) +  # Add horizontal dashed line at y = 1
  scale_color_manual(values = color_mapping) +
  labs(title = "",
       x = "Position of mutation of HD alignment",
       y = "",
       color = "Annotation according to Kock et al.") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),  # Remove background grid
    axis.text.x = element_text(angle = 90, vjust = 0.5)
  )

wilcox.test(data$trueMagnitude[data$color_category != "Mutation has no effect"], data$trueMagnitude[data$color_category == "Mutation has no effect"])
```

```{r S10B}
# Create a data frame for plotting
data <- data.frame(
  position = mutationInfo$position[mutantID]-1,
  R2List = R2ListbaseLine,
  trueMagnitude = dddGRangeTrue
)

# Define the color categories
data$color_category <- ifelse(
  ((matchSampleInfo$affinity_change == 'no change') + (matchSampleInfo$specificity_change == 'no')) == 2, "Mutation has no effect", "Specificity or affinity change")

# Define color mapping
color_mapping <- c("Specificity or affinity change" = "red",  "Mutation has no effect" = "cyan")

# First Lollipop Plot: R2List
ggplot(data, aes(x = factor(position), y = R2List, color = color_category)) +
  geom_segment(aes(xend = factor(position), y = 0, yend = R2List), 
               color = "gray", linetype = "dotted", size = 0.5) +  # Gray, dotted, thin sticks
  geom_point(size = 3, alpha = 1) +  
  scale_color_manual(values = color_mapping) +
  labs(title = "",
       x = "Position of mutation of HD alignment",
       y = "",
       color = "Annotation according to Kock et al.") +
  theme_minimal() +
  theme(
    panel.grid = element_blank(),  # Remove background grid
    axis.text.x = element_text(angle = 90, vjust = 0.5)
  )
```

# Figure S11
```{r S11A-L}
mutantID <- which(mutationInfo$mutate != 'WT')
WTID <- which(mutationInfo$mutate == 'WT')

standard_Alignment <- list()
for(i in 1:nrow(Kock_alignment)){
  WTid <- which(CV_alignment$name == mutationInfo$TF[i])
  add <- CV_alignment[WTid[1],]
  standard_Alignment <- rbind.data.frame(standard_Alignment, add)
}

#WTmodel <- makeSVDModel(CV_motif_Kock, CV_alignment_Kock, Positions = motifPoses, Ftest_pVal = 0.01)

mutationPredList <- list()
mutationPredConfidence <- c()
inTraining <- c()
for(i in mutantID){
  mutpos <- as.numeric(mutationInfo$position[i])-1
  svdModel <- makeSVDModel.singleMutation(CV_motif_Kock, CV_alignment_Kock, Positions = motifPoses, mutation = mutpos)
  inTraining <- c(sum(substr(Kock_alignment$alignment[i],mutpos,mutpos) == substr(CV_alignment_Kock$alignment,mutpos,mutpos)), inTraining)
  #predict mutant
  predPSAM <- data.frame(pred = c(0,0,0,0))
  rownames(predPSAM) <- DNA()
  for(pos in motifPoses){
    predColumn <- predict(svdModel[[pos]], Kock_alignment[i,], useSimilarAA = T)
    predPSAM <- cbind.data.frame(predPSAM, predColumn)
    predConfidence <- attr(predColumn, 'confidence')
  }
  predPSAM <- predPSAM[,-1]
  colnames(predPSAM) <- motifPoses
  #predict standard
  
  mutationPredList[[length(mutationPredList)+1]] <- list(name = Kock_alignment$name[i], matrix = predPSAM)
  mutationPredConfidence <- c(mutationPredConfidence, predConfidence)
}


predMotifs <- mutationPredList
sum(unlist(lapply(predMotifs, function(x) x$name)) == unlist(lapply(trueMotifs, function(x) x$name)))

dddGTrue <- list()
dddGpred <- list()
R2List <- c()
for(i in 1:length(trueMotifs)){
  predDiffTetra <- matrix2tetrahedron(predMotifs[[i]]$matrix) - matrix2tetrahedron(trueStandard[[i]]$matrix)
  if(sum(predDiffTetra) == 0){
    predDiffTetra[1] <- 0.01
    mod <- TRUE
  }else{
    mod <- FALSE
  }
  trueDiffTetra <- matrix2tetrahedron(trueMotifs[[i]]$matrix) - matrix2tetrahedron(trueStandard[[i]]$matrix)
  predDiffMatrix <- tetrahedron2matrix(predDiffTetra)
  trueDiffMatrix <- tetrahedron2matrix(trueDiffTetra)
  predDiffMatrix[predDiffMatrix < 0.01] <- 0.01
  trueDiffMatrix[trueDiffMatrix < 0.01] <- 0.01
  rownames(trueDiffMatrix) <- DNA()
  rownames(predDiffMatrix) <- DNA()
  if(mod){
    predDiffMatrix[1,] <- 1
  }
  pred <- as.numeric(frequency2ddG(predDiffMatrix))
  true <- as.numeric(frequency2ddG(trueDiffMatrix))
  predTrue <- data.frame(true = true, pred = pred)
  dddGTrue[[length(dddGTrue) + 1]] <- trueDiffMatrix
  dddGpred[[length(dddGpred) + 1]] <- predDiffMatrix
  #plotPredTrue(predTrue, main = paste0('FamilyCode rediction for ',unlist(lapply(trueMotifs, function(x) x$name))[i]), xlab = 'Experimental -ΔΔΔG/RT', ylab = 'Predicted -ΔΔΔG/RT')
  R2List <- c(R2List, groupedR2(predTrue$true, predTrue$pred, member = nrow(predTrue)))
}

ids <- c(49,58,50,52)
for(i in ids){
  predDiffTetra <- matrix2tetrahedron(predMotifs[[i]]$matrix) - matrix2tetrahedron(trueStandard[[i]]$matrix)
  if(sum(predDiffTetra) == 0){
    predDiffTetra[1] <- 0.01
    mod <- TRUE
  }else{
    mod <- FALSE
  }
  trueDiffTetra <- matrix2tetrahedron(trueMotifs[[i]]$matrix) - matrix2tetrahedron(trueStandard[[i]]$matrix)
  predDiffMatrix <- tetrahedron2matrix(predDiffTetra)
  trueDiffMatrix <- tetrahedron2matrix(trueDiffTetra)
  predDiffMatrix[predDiffMatrix < 0.01] <- 0.01
  trueDiffMatrix[trueDiffMatrix < 0.01] <- 0.01
  rownames(trueDiffMatrix) <- DNA()
  rownames(predDiffMatrix) <- DNA()
  if(mod){
    predDiffMatrix[1,] <- 1
  }
  pred <- as.numeric(frequency2ddG(predDiffMatrix))
  true <- as.numeric(frequency2ddG(trueDiffMatrix))
  predTrue <- data.frame(true = true, pred = pred)
  plotPredTrue(predTrue, main = paste0('FamilyCode prediction for ',unlist(lapply(trueMotifs, function(x) x$name))[i]), xlab = 'Replicate1 -ΔΔΔG/RT', ylab = 'Predicted -ΔΔΔG/RT')
  plot(mononucleotide_logo(frequency2ddG(predDiffMatrix), label = F))
  plot(mononucleotide_logo(frequency2ddG(trueDiffMatrix), label = F))
  #mononucleotide_logo(frequency2ddG(Kock_motifs_set[[mut]]$matrix))
  #mononucleotide_logo(frequency2ddG(Kock_motifs_set[[max(WTID[WTID < mut])]]$matrix))
  
  mutid <- mutantID[i]
  predDiffTetra <- matrix2tetrahedron(Exp2TrueMotifs[[mutid]]$matrix) - matrix2tetrahedron(trueStandard[[i]]$matrix)
  #predDiffTetra <- matrix2tetrahedron(Exp2TrueMotifs[[mutid]]$matrix) - matrix2tetrahedron(Exp2TrueMotifs[[max(WTID[WTID<mutid])]]$matrix)
  trueDiffTetra <- matrix2tetrahedron(trueMotifs[[i]]$matrix) - matrix2tetrahedron(trueStandard[[i]]$matrix)
  #trueDiffTetra <- matrix2tetrahedron(trueMotifs[[i]]$matrix) - matrix2tetrahedron(Exp2TrueMotifs[[max(WTID[WTID<mutid])]]$matrix)
  predDiffMatrix <- tetrahedron2matrix(predDiffTetra)
  trueDiffMatrix <- tetrahedron2matrix(trueDiffTetra)
  predDiffMatrix[predDiffMatrix < 0.01] <- 0.01
  trueDiffMatrix[trueDiffMatrix < 0.01] <- 0.01
  rownames(trueDiffMatrix) <- DNA()
  rownames(predDiffMatrix) <- DNA()
  pred <- as.numeric(frequency2ddG(predDiffMatrix))
  true <- as.numeric(frequency2ddG(trueDiffMatrix))
  predTrue <- data.frame(true = true, pred = pred)
  plotPredTrue(predTrue, main = paste0('Base Line for ',unlist(lapply(trueMotifs, function(x) x$name))[i]), xlab = 'Replicate1 -ΔΔΔG/RT', ylab = 'Replicate2 -ΔΔΔG/RT')
  plot(mononucleotide_logo(frequency2ddG(predDiffMatrix), label = F))
}



```

