---
title: "FamilyCode on CisBP data (bHLH data with De masi et. al., comparison)"
output: html_notebook
---

## Family Code application on combinatory data in CisBP. Using bHLH proteins as an example
### Import library and data
```{r}
library(FamilyCode)
workDir <- 'cisBP_bHLH/'
#download from cisBP 'bulk download' tab, incuding pwms and TF_Information
TF_Information <- read.delim("cisBP_bHLH/TF_Information.txt")
TFDirect <- TF_Information[TF_Information$TF_Status == 'D',]
#filter out methylated data sets
TFDirect <- TFDirect[-grep('Methyl', TFDirect$DBID.1),]
#filter out demasi mutation sets
TFdemasi <- TFDirect[TFDirect$MSource_Identifier == 'DeMasi2011',]
TFDirect <- TFDirect[TFDirect$MSource_Identifier != 'DeMasi2011',]
```

### download Protein data from Uniprot
```{r}
#uniprot can only handle certain number of searches at a time, so separate all data into two parts
gene_ids <- unique(TFDirect[,c('TF_Name', 'TF_Species')])
searchUnitTemplate <- '(gene:"GENE" AND "SPECIES")'
searchQ <- gsub('GENE', gene_ids$TF_Name[1], searchUnitTemplate)
searchQ <- gsub('SPECIES', gene_ids$TF_Species[1], searchQ)
for(i in 1:140){#nrow(gene_ids)
  addQ <- gsub('GENE', gene_ids$TF_Name[i], searchUnitTemplate)
  addQ <- gsub('SPECIES', gene_ids$TF_Species[i], addQ)
  searchQ <- paste0(searchQ, ' OR ', addQ)
}
cat(searchQ)
```

```{r}
searchQ <- gsub('GENE', gene_ids$TF_Name[1], searchUnitTemplate)
searchQ <- gsub('SPECIES', gene_ids$TF_Species[1], searchQ)
for(i in 141:nrow(gene_ids)){#nrow(gene_ids)
  addQ <- gsub('GENE', gene_ids$TF_Name[i], searchUnitTemplate)
  addQ <- gsub('SPECIES', gene_ids$TF_Species[i], addQ)
  searchQ <- paste0(searchQ, ' OR ', addQ)
}
cat(searchQ)
```

### Import protein data 
```{r}
geneInfoUniprot1 <- read.delim("cisBP_bHLH/geneInfoUniprot_all1.tsv")
geneInfoUniprot2 <- read.delim("cisBP_bHLH/geneInfoUniprot_all2.tsv")
geneInfoUniprot <- rbind.data.frame(geneInfoUniprot1, geneInfoUniprot2)
geneInfoUniprot$Gene.Names <- paste0(geneInfoUniprot$Gene.Names, ' ')
```

### Match protein sequences data from uniprot with motif information from cisBP
```{r}

geneInfo <- data.frame()
for(i in 1:nrow(gene_ids)){
  add <- geneInfoUniprot[grep(paste0(gene_ids$TF_Name[i],' '), geneInfoUniprot$Gene.Names),]
  if(nrow(add) != 0){
    add <- add[nchar(add$Sequence) == max(nchar(add$Sequence)),][1,]
    addSeq <- add$Sequence
  }else{
    addSeq <- NA 
  }
  Motif <- unique(TFDirect[TFDirect$TF_Name == gene_ids$TF_Name[i],]$Motif_ID)
  addLine <- data.frame(name = gene_ids$TF_Name[i], seq = addSeq, motif = Motif)
  geneInfo <- rbind.data.frame(geneInfo, addLine)
}

geneInfo <- geneInfo[!is.na(geneInfo$seq),]
#write sequences as fasta to
writeFasta(geneInfo, 'cisBP_bHLH/TF.fa')
```

### Use hmmalign to align the output sequences with the Homeo Domain HMM model from pfam entry PF00010
```{}
hmmalign -o cisBP_TF.sto bHLH.hmm TF.fa
```

### Use Jalview to hide non-hmm columns and output filtered residues, saved as TFhmm.sto
### Screen and Trim unusable alignment entries
```{r}
ali <- read.table(r"(cisBP_bHLH/TFhmm.sto)", quote="\"", fill = T)
ali <- ali[-nrow(ali),]
ali$V1 <- apply(ali, 1, function(x) strsplit(x[1],'/')[[1]][1])
ali$V1 <- apply(ali, 1, function(x) strsplit(x[1],'\\|')[[1]][2])
colnames(ali) <- c('name','alignment')
keep <- which(substr(ali$alignment, 9,9) == 'E')
cisBP_alignment <- ali[keep,]
geneInfo <- geneInfo[keep,]
#Check if cisBP_alignment and geneInfo matches
sum(cisBP_alignment$name != geneInfo$name) == 0
```

### Loading pwms from cisBP using geneInfo as index
```{r}
cisBP_motifs <- list()
for(i in 1:nrow(geneInfo)){
  motif <- geneInfo$motif[i]
  motif <- t(read.table(paste0('cisBP_bHLH/pwms/', motif, '.txt'), header = T, row.names = 1))
  motif[motif < 0.001] <- 0.001#add psuedo count
  add <- list()
  add$gene_symbol <- geneInfo$name[i]
  add$study <- i
  add$mode <- 1
  add$matrix <- as.matrix(motif)
  cisBP_motifs[[length(cisBP_motifs)+1]] <- add
  add <- list()
  #add reverse compliment motif
  add$gene_symbol <- geneInfo$name[i]
  add$study <- i
  add$mode <- -1
  if(ncol(motif) == 0){
    rev <- motif
  }else{
    rev <- as.matrix(motif)[nrow(motif):1, ncol(motif):1]
  }
  rownames(rev) <- DNA()
  add$matrix <- rev
  cisBP_motifs[[length(cisBP_motifs)+1]] <- add
}
#check if length of cisBP_motif is equal to 2 times geneInfo
length(cisBP_motifs) == nrow(geneInfo)*2
```
### Load bHLH 6mer CANNTG and score motifs with the seed
```{r}
#align to CANNTG motif
bHLHseed <- matrix(nrow = 4,ncol = 6,data = c(0,1,0,0,
                                              1,0,0,0,
                                              0,0,0,0,
                                              0,0,0,0,
                                              0,0,0,1,
                                              0,0,1,0
))
rownames(bHLHseed) <- DNA()
#Scoring would result in some motif shorter than the seed, those motifs will be discarded. 
motifScores <- scoreMotifList(cisBP_motifs, bHLHseed, weight = c(1,1,0,0,1,1))
```

### Filter and reorder motif information according to motif matching score.
The motif (forward or reverse compliment) with the lowest motif matching score of each entry will be kept. For the same protein, entries are arranged from lower score to higher score
```{r}
idstds <- unique(motifScores[,1:2])
all_motifs <- data.frame(NULL)
for(i in 1:nrow(idstds)){
  fDT <- motifScores[motifScores$gene_symbol == idstds[i,1],]
  fDT <- fDT[fDT$study == idstds[i,2],]
  all_motifs <- rbind.data.frame(all_motifs, fDT[which.min(fDT$score),])
}
motifs_arranged <- data.frame(NULL)
motifScore <- c()
for(i in 1:length(unique(all_motifs$gene_symbol))){
  add <- all_motifs[all_motifs$gene_symbol == unique(all_motifs$gene_symbol)[i],]
  add <- dplyr::arrange(add, score)
  motifScore <- c(motifScore,add$score[1])
  motifs_arranged <- rbind.data.frame(motifs_arranged, add)
}
#Plot distribution of lowest motif matching score
plot.ecdf(as.numeric(motifScore))
```

```{r}
#Filter out motifs that do not match well with the seed
motifTable_baseline <- motifs_arranged[motifs_arranged$score < 0.2,]
#Number of unique HD protiens kept:
length(unique(motifTable_baseline$gene_symbol))
```

### Construct motif list and alignment table for Base Line (Motif similarity between the two best experiments)
```{r}
motifPoses <- c('P-3','P-2','P-1','P1','P2','P3')
cisBP_motifs_baseline <- filterMotifList(motifTable_baseline, cisBP_motifs, 6, motifPoses)
cisBP_alignment <- cisBP_alignment[match(motifTable_baseline$gene_symbol, cisBP_alignment$name),]
#check if motif List has same length as alignment
length(cisBP_motifs_baseline) == nrow(cisBP_alignment)
```
### Show base line accuracy
```{r}
for(pos in motifPoses){
  ScreenPoses <- c(pos,pos)
  BaseLines.ddG <- getBaseLineAccuracy(cisBP_motifs_baseline, pos = ScreenPoses, randomSample = F)
  colnames(BaseLines.ddG) <- c('pred','true')
  BaseLines.PFM <- predTrue.ddG2frequency(BaseLines.ddG, PFM = T)
  plotPredTrue(BaseLines.ddG, xlab = 'Experiment 1', ylab = 'Experiment 2', main = paste0('Pair-wise Experiment comparison at ',pos))
  boxplot(groupedR2(BaseLines.ddG$pred, BaseLines.ddG$true, member = length(ScreenPoses)*4, mean= F), ylim = c(0,1),
        main = paste0(pos,' Mean R^2 = ', round(groupedR2(BaseLines.ddG$pred, BaseLines.ddG$true, 
                                                          member = length(ScreenPoses)*4, mean= T),4)))
  plotPredTrue(BaseLines.PFM, xlab = 'Experiment 1', ylab = 'Experiment 2', main = paste0('Base Line PFM at ',pos))
  hit <- sum(round(groupedR2(BaseLines.PFM$pred, BaseLines.PFM$true, member = 4, mean= F)^(1/2),4) > 0.5)/(nrow(BaseLines.PFM)/4)
  boxplot(groupedR2(BaseLines.PFM$pred, BaseLines.PFM$true, member = length(ScreenPoses)*4, mean= F)^(1/2), ylim = c(0,1),
        main = paste0(pos, ' Mean PCC = ', round(groupedR2(BaseLines.PFM$pred, BaseLines.PFM$true, 
                                                     member = length(ScreenPoses)*4, mean= T)^(1/2),4), ' Accuracy = ', round(hit,4)))
}

```

### Make p-val table
```{r}
del <- which(duplicated(cisBP_alignment$alignment))
CV_alignment <- cisBP_alignment[-del,]
CV_motif <- cisBP_motifs_baseline[-del]
#genreate p-val tablle
pvalTable <- getPvalTable(CV_motif, Alignment = CV_alignment, pos_index = motifPoses)
#apply -log(10)
pvalTable <- -log(pvalTable,10)
#set NAs to 0
pvalTable[is.na(pvalTable)] <- 0
#plot heatmap
gplots::heatmap.2(as.matrix(as.data.frame(lapply(pvalTable, as.numeric))),dendrogram='none',
                  Rowv=FALSE, Colv=FALSE,trace='none',col = rev(heat.colors(12)), key.title = '-logP-val')
```

### Test on FamilyCode prediction with leave-one-out CV
Compare with paper: Wetzel, J. L., Zhang, K., & Singh, M. (2022). Learning probabilistic protein–DNA recognition codes from DNA-binding specificities using structural mappings. Genome Research, 32(9), 1776-1786.
Accuracy measurement in Figure 2
"PWM F, where entry fbj is the frequency with which nucleotide b is observed in column j in a set of aligned binding sites of that TF. "
```{r}
FCpredTrueList <- list()
for(pos in motifPoses){
  predTrue <- SVDregression.Iterative.CV(CV_motif,CV_alignment, pos = pos,Ftest_pVal = 0.001, useSimilarAA = T)
  #delete unsuccessful predictions
  predTrue <- predTrue[!is.na(predTrue$pred),]
  FCpredTrueList[[pos]] <- predTrue
  print(pos)
  
}
for(pos in motifPoses){
  predTrue <- FCpredTrueList[[pos]]
  #delete unsuccessful predictions
  predTrue <- predTrue[!is.na(predTrue$pred),]
  
  predTrue.PFM <- predTrue.ddG2frequency(predTrue, PFM = T)
  plotPredTrue(predTrue, main = paste0('FamilyCode at ',pos))
  boxplot(groupedR2(predTrue$pred, predTrue$true, member = 4, mean= F), ylim = c(0,1),
        main = paste0(pos,' Mean R^2 = ', round(groupedR2(predTrue$pred, predTrue$true, 
                                                          member = 4, mean= T),4)))
  #plot confidance
  attr(predTrue, 'confidence') <- attr(predTrue, 'confidence')[attr(predTrue, 'confidence') <= 1]
  plot(attr(predTrue, 'confidence'),groupedR2(predTrue$pred, predTrue$true,member = 4, mean = F), ylab = 'Predict-True R^2', xlab = 'Confidence score', main = paste0('Confidence at ', pos))
  legend('bottomright', paste0('R^2 = ', round(groupedR2(groupedR2(predTrue$pred, predTrue$true,member = 4, mean = F),
                                                     attr(predTrue, 'confidence'), ),4)), bty = 'n')
  dt <- data.frame(conf = attr(predTrue, 'confidence'), R2 = groupedR2(predTrue$pred, predTrue$true,member = 4, mean = F))
  lm <- lm(R2~conf, data= dt)
  abline(lm, col= 'red')
  
  plotPredTrue(predTrue.PFM, main = paste0('FamilyCode prediction PFM at ',pos))
  hit <- sum(round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F)^(1/2),4) > 0.5)/(nrow(predTrue.PFM)/4)
  boxplot(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F)^(1/2), ylim = c(0,1),
        main = paste0(pos, ' Mean PCC = ', round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, 
                                                     member = 4, mean= T)^(1/2),4), ' Accuracy = ', round(hit,4)))
}
```
### Comparing prediction result with Closest sequence
from paper:
Weirauch, M. T., Yang, A., Albu, M., Cote, A. G., Montenegro-Montero, A., Drewe, P., ... & Hughes, T. R. (2014). Determination and inference of eukaryotic transcription factor sequence specificity. Cell, 158(6), 1431-1443.
Figure S6/S7
```{r}
for(pos in motifPoses){
  del <- which(duplicated(cisBP_alignment$alignment))
  CV_alignment <- cisBP_alignment[-del,]
  CV_motif <- cisBP_motifs_baseline[-del]
  predTrue <- closestSeqPred(CV_motif,CV_alignment, pos = pos)
  #delete unsuccessful predictions
  predTrue <- predTrue[!is.na(predTrue$pred),]
  predTrue.PFM <- predTrue.ddG2frequency(predTrue, PFM = T)
  
  plot.ecdf(predTrue$similarity)
  plotPredTrue(predTrue, main = paste0('Closest Sequence prediction at ',pos))
  boxplot(groupedR2(predTrue$pred, predTrue$true, member = 4, mean= F), ylim = c(0,1),
        main = paste0(pos,' Mean R^2 = ', round(groupedR2(predTrue$pred, predTrue$true, 
                                                          member = 4, mean= T),4)))
  
  plotPredTrue(predTrue.PFM, main = paste0('Closest sequence prediction PFM at ',pos))
  hit <- sum(round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F)^(1/2),4) > 0.5)/(nrow(predTrue.PFM)/4)
  boxplot(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F)^(1/2), ylim = c(0,1),
        main = paste0(pos, 'Mean PCC = ', round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, 
                                                     member = 4, mean= T)^(1/2),4), ' Accuracy = ', round(hit,4)))
  
}
```
### Comparing prediction result with Similarity regression
from paper:
Lambert, S. A., Yang, A. W., Sasse, A., Cowley, G., Albu, M., Caddick, M. X., ... & Hughes, T. R. (2019). Similarity regression predicts evolution of transcription factor sequence specificity. Nature genetics, 51(6), 981-989.
Figure S6/S7

```{r}
for(pos in motifPoses){
  del <- which(duplicated(cisBP_alignment$alignment))
  CV_alignment <- cisBP_alignment[-del,]
  mod_alignment <- CV_alignment
  mod_alignment$alignment <- paste0(substr(mod_alignment$alignment,1,34), '-','-',substr(mod_alignment$alignment,35,53))
  CV_motif <- cisBP_motifs_baseline[-del]
  predTrue <- SRpred(CV_motif,mod_alignment, pos = pos, weightfile = 'cisBP_bHLH/F082_1.97d.json')
  #delete unsuccessful predictions
  predTrue <- predTrue[!is.na(predTrue$pred),]
  predTrue.PFM <- predTrue.ddG2frequency(predTrue, PFM = T)
  
  plot.ecdf(predTrue$similarity)
  plotPredTrue(predTrue, main = paste0('Similarity regression at ',pos))
  boxplot(groupedR2(predTrue$pred, predTrue$true, member = 4, mean= F), ylim = c(0,1),
        main = paste0(pos,' Mean R^2 = ', round(groupedR2(predTrue$pred, predTrue$true, 
                                                          member = 4, mean= T),4)))
  
  plotPredTrue(predTrue.PFM, main = paste0('Similarity regression prediction PFM at ',pos))
  hit <- sum(round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F)^(1/2),4) > 0.5)/(nrow(predTrue.PFM)/4)
  boxplot(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F)^(1/2), ylim = c(0,1),
        main = paste0(pos, 'Mean PCC = ', round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, 
                                                     member = 4, mean= T)^(1/2),4), ' Accuracy = ', round(hit,4)))
}

```

###Train model with all entries of cisBP bHLH data
```{r}
del <- which(duplicated(cisBP_alignment$alignment))
CV_alignment <- cisBP_alignment[-del,]
CV_motif <- cisBP_motifs_baseline[-del]
svdModel <- makeSVDModel(CV_motif, CV_alignment, Positions = c('P-3','P-2','P-1','P1','P2','P3'))
```

### Load De Masi motifs
```{r}
DeMasiInfo <- data.frame(name = TFdemasi$DBID.1, seq = CV_alignment[CV_alignment$name == 'hlh-1',]$alignment, motif = TFdemasi$Motif_ID)
substr(DeMasiInfo$seq,13,13) <- c('R','T','V','L')

DeMasi_motifs <- list()
for(i in 1:nrow(DeMasiInfo)){
  motif <- DeMasiInfo$motif[i]
  motif <- t(read.table(paste0('cisBP_bHLH/pwms/', motif, '.txt'), header = T, row.names = 1))
  motif[motif < 0.001] <- 0.001#add psuedo count
  add <- list()
  add$gene_symbol <- DeMasiInfo$name[i]
  add$study <- i
  add$mode <- 1
  add$matrix <- as.matrix(motif)
  DeMasi_motifs[[length(DeMasi_motifs)+1]] <- add
  add <- list()
  #add reverse compliment motif
  add$gene_symbol <- DeMasiInfo$name[i]
  add$study <- i
  add$mode <- -1
  if(ncol(motif) == 0){
    rev <- motif
  }else{
    rev <- as.matrix(motif)[nrow(motif):1, ncol(motif):1]
  }
  rownames(rev) <- DNA()
  add$matrix <- rev
  DeMasi_motifs[[length(DeMasi_motifs)+1]] <- add
}

#Scoring would result in some motif shorter than the seed, those motifs will be discarded. 
DeMasimotifScores <- scoreMotifList(DeMasi_motifs, bHLHseed, weight = c(1,1,0,0,1,1))

idstds <- unique(DeMasimotifScores[,1:2])
Demasi_motifsDT <- data.frame(NULL)
for(i in 1:nrow(idstds)){
  fDT <- DeMasimotifScores[DeMasimotifScores$gene_symbol == idstds[i,1],]
  fDT <- fDT[fDT$study == idstds[i,2],]
  Demasi_motifsDT <- rbind.data.frame(Demasi_motifsDT, fDT[which.min(fDT$score),])
}

DeMasi_motifs <- filterMotifList(Demasi_motifsDT, DeMasi_motifs, 6, motifPoses)
```

### DeMasi Baseline
```{r}
PBM <- CV_motif[CV_alignment$name == 'hlh-1'][[1]]$matrix
Demasi <- DeMasi_motifs[[4]]$matrix
true <- frequency2ddG(PBM[,c('P-1','P1')])

pred <- frequency2ddG(Demasi[,c('P-1','P1')])


predTrueBaseline <- data.frame(pred = as.numeric(unlist(pred)), true = as.numeric(unlist(true)))
plotPredTrue(predTrueBaseline, xlab = 'Grove et al. (2009)', ylab = 'DeMasi et al. (2011)', main = paste0('Pair-wise experiment comparison at P-/+1'))

```

### Predict for Demasi motifs
```{r}
pred <- lapply(DeMasiInfo$seq, function(x) SVDmodelPredPSAM(svdModel[c('P-1','P1')], x))
pred <- lapply(pred, frequency2ddG)
true <- lapply(DeMasi_motifs, function(x) x$matrix[,c('P-1','P1')])
true <- lapply(true, frequency2ddG)
predTrueDeMasi <- data.frame(pred = as.numeric(unlist(pred)), true = as.numeric(unlist(true)))
plotPredTrue(predTrueDeMasi[1:24,])
plotPredTrue(predTrueDeMasi[1:24,], main = paste0('FamilyCode prediction at P-/+1'))
```

### Generate prediction motif models
```{r}
pos = c('P-3','P-2','P-1','P1','P2','P3')
demasi.pred <- lapply(DeMasiInfo$seq, function(x) SVDmodelPredPSAM(svdModel[pos], x))

for(i in 1:length(demasi.pred)){
  plot <- mononucleotide_logo(frequency2ddG(DeMasi_motifs[[i]]$matrix))
  plot <- plot + ggplot2::ggtitle(paste0(DeMasiInfo$name[i],'_true'))
  plot(plot)
  plot <- mononucleotide_logo(frequency2ddG(demasi.pred[[i]]))
  plot <- plot + ggplot2::ggtitle(paste0(DeMasiInfo$name[i], '_pred'))
  plot(plot)
}
```

### Similarity regression for De masi
```{r}
mod_alignment <- CV_alignment
mod_alignment$alignment <- paste0(substr(mod_alignment$alignment,1,34), '-','-',substr(mod_alignment$alignment,35,53))

DeMasiAli <- data.frame(name = DeMasiInfo$name, alignment = paste0(substr(DeMasiInfo$seq,1,34), '-','-',substr(DeMasiInfo$seq,35,53)))
pred_motifs <- SRpredPSAM(DeMasiAli, CV_motif, mod_alignment, weightfile = 'cisBP_bHLH/F082_1.97d.json')

pred <- lapply(pred_motifs, function(x) x$matrix[,c('P-1','P1')])
pred <- lapply(pred, frequency2ddG)
true <- lapply(DeMasi_motifs, function(x) x$matrix[,c('P-1','P1')])
true <- lapply(true, frequency2ddG)
predTrueDeMasi <- data.frame(pred = as.numeric(unlist(pred)), true = as.numeric(unlist(true)))
plotPredTrue(predTrueDeMasi[1:24,], main = paste0('Similarity regression at P-/+1'))
```
### Closest sequence for De masi
```{r}
pred_motifs <- CSpredPSAM(DeMasiAli, CV_motif, CV_alignment)

pred <- lapply(pred_motifs, function(x) x$matrix[,c('P-1','P1')])
pred <- lapply(pred, frequency2ddG)
true <- lapply(DeMasi_motifs, function(x) x$matrix[,c('P-1','P1')])
true <- lapply(true, frequency2ddG)
predTrueDeMasi <- data.frame(pred = as.numeric(unlist(pred)), true = as.numeric(unlist(true)))
plotPredTrue(predTrueDeMasi[1:24,], main = paste0('Closest sequence prediction at P-/+1'))
```


### systemetic test of prediction methods with subsampling
```{r}
options(warn=0)
del <- which(duplicated(cisBP_alignment$alignment))
CV_alignment <- cisBP_alignment[-del,]
CV_motif <- cisBP_motifs_baseline[-del]
nset <- nrow(CV_alignment)
results <- list()
for(i in 2:10){
  FCR2List <- c()
  FCrmsdList <- c()
  CSR2List <- c()
  CSrmsdList <- c()
  SRR2List <- c()
  SRrmsdList <- c()
  for(j in 1:40){
    trainID <- sample(1:nset, floor(nset-nset/(i)))
    train_alignment <- CV_alignment[trainID,]
    train_motifs <- CV_motif[trainID]
    test_alignment <- CV_alignment[-trainID,]
    test_motifs <- CV_motif[-trainID]
    #FC
    svdModel <- makeSVDModel(train_motifs, train_alignment, Positions = c('P-3','P-2','P-1','P1','P2','P3'))
    pred <- lapply(test_alignment$alignment, function(x) SVDmodelPredPSAM(svdModel[c('P-1','P1')], x))
    pred <- lapply(pred, frequency2ddG)
    true <- lapply(test_motifs, function(x) x$matrix[,c('P-1','P1')])
    true <- lapply(true, frequency2ddG)
    predTrue <- data.frame(pred = as.numeric(unlist(pred)), true = as.numeric(unlist(true)))
    FCR2 <- round(groupedR2(predTrue$true, predTrue$pred),4)
    FCrmsd <- round(RMSD(predTrue$true,predTrue$pred),4)
    #CS
    pred_motifs <- CSpredPSAM(test_alignment, train_motifs, train_alignment)
    pred <- lapply(pred_motifs, function(x) x$matrix[,c('P-1','P1')])
    pred <- lapply(pred, frequency2ddG)
    true <- lapply(test_motifs, function(x) x$matrix[,c('P-1','P1')])
    true <- lapply(true, frequency2ddG)
    predTrue <- data.frame(pred = as.numeric(unlist(pred)), true = as.numeric(unlist(true)))
    CSR2 <- round(groupedR2(predTrue$true, predTrue$pred),4)
    CSrmsd <- round(RMSD(predTrue$true,predTrue$pred),4)
    #SR
    
    train_alignment$alignment <- paste0(substr(train_alignment$alignment,1,34), '-','-',substr(train_alignment$alignment,35,53))
    test_alignment$alignment <- paste0(substr(test_alignment$alignment,1,34), '-','-',substr(test_alignment$alignment,35,53))
    pred_motifs <- SRpredPSAM(test_alignment, train_motifs, train_alignment, weightfile = 'cisBP_bHLH/F082_1.97d.json')
    pred <- lapply(pred_motifs, function(x) x$matrix[,c('P-1','P1')])
    pred <- lapply(pred, frequency2ddG)
    true <- lapply(test_motifs, function(x) x$matrix[,c('P-1','P1')])
    true <- lapply(true, frequency2ddG)
    predTrue <- data.frame(pred = as.numeric(unlist(pred)), true = as.numeric(unlist(true)))
    SRR2 <- round(groupedR2(predTrue$true, predTrue$pred),4)
    SRrmsd <- round(RMSD(predTrue$true,predTrue$pred),4)
    FCR2List <- c(FCR2List, FCR2)
    FCrmsdList <- c(FCrmsdList, FCrmsd)
    CSR2List <- c(CSR2List, CSR2)
    CSrmsdList <- c(CSrmsdList, CSrmsd)
    SRR2List <- c(SRR2List, SRR2)
    SRrmsdList <- c(SRrmsdList, SRrmsd)
  }
  add <- list()
  add$FCR2 <- FCR2List
  add$FCrmsd <- FCrmsdList
  add$CSR2 <- CSR2List
  add$CSrmsd <- CSrmsdList
  add$SRR2 <- SRR2List
  add$SRrmsd <- SRrmsdList
  results[[length(results)+1]] <- add
}
bHLH_cisBP_results <- results
FCR2 <- unlist(lapply(results, function(x) mean(x$FCR2)))
FCrmsd <- unlist(lapply(results, function(x) mean(x$FCrmsd)))
CSR2 <- unlist(lapply(results, function(x) mean(x$CSR2)))
CSrmsd <- unlist(lapply(results, function(x) mean(x$CSrmsd)))
SRR2 <- unlist(lapply(results, function(x) mean(x$SRR2)))
SRrmsd <- unlist(lapply(results, function(x) mean(x$SRrmsd)))

matplot(1-1/2:10, cbind(FCR2, CSR2, SRR2), type = "l", lty = 1, 
        col = c("red", "blue", "green"), xlab = "X", 
        ylab = "Y", main = "Multiple Lines Plot")
matplot(1-1/2:10, cbind(FCrmsd, CSrmsd, SRrmsd), type = "l", lty = 1, 
        col = c("red", "blue", "green"), xlab = "X", 
        ylab = "Y", main = "Multiple Lines Plot")
```

### score PBM with cisBP motif with probound
```{r}
mkdirs(paste0(workDir,'JSONs'))
CTs <- list.files(paste0(workDir,'CountTables/'))
library(jsonlite)
sampleJSON <- 'sample.json'
json <- fromJSON(sampleJSON)
json$modelSettings$bindingModes$size <- 6
for(i in 1:nrow(CV_alignment)){
  testID <- CV_alignment$name[i]
  motifID <- TFDirect$Motif_ID[TFDirect$TF_Name == testID]
  if(length(motifID) == 0){
    next()
  }
  if(length(which(paste0(motifID[1],'_CT.tsv') == CTs)) == 0){
    next()
  }
  pwm <- apply(frequency2ddG(CV_motif[[i]]$matrix), 2, function(x) x-max(x))
  json$coefficients$bindingModes$mononucleotide <- list(as.numeric(pwm))
  outJSON <- paste0(workDir,'JSONs/',motifID[1],'.json')
  write(toJSON(json, pretty = TRUE), outJSON)
}
```

```{r}
PBpwmPCC <- read.table(paste0(workDir,"PBpwmPCC.txt"), quote="\"", comment.char="")
median(PBpwmPCC$V1)
median(pbPCC$V1)
cisBPPCC <- PBpwmPCC
```
