---
title: "FamilyCode on CisBP data Processing PBM data with pyProBound (HD data)"
output: html_notebook
---

## Family Code application on combinatory data in CisBP. Using HD proteins as an example
### Import library and data
```{r}
library(FamilyCode)
library(R.utils)
workDir <- 'intermediateData/HD/PBM_HD/'
rawData <- 'rawData/HD/'
#download from cisBP 'bulk download' tab, inducing Z-score and TF_Information
TF_Information <- read.delim(paste0(rawData, "TF_Information.txt"))
Zscores <- read.delim(paste0(rawData,"Zscores.txt"))
#filter for PBM data with Z-scores
TFDirect <- TF_Information[TF_Information$TF_Status == 'D',]
rawPBM <- substr(colnames(Zscores)[-1], 1, 11)
#deal with duplicated experiment
substr(rawPBM[duplicated(rawPBM)],11,11) <- '1'
colnames(Zscores)[-1] <- rawPBM
TFDirect.duplicated <- TFDirect
substr(TFDirect.duplicated$Motif_ID,11,11) <- '1'
TFDirect <- rbind.data.frame(TFDirect, TFDirect.duplicated)
#get cisBP entries with PBM Z-scores
TFDirect <- TFDirect[match(intersect(TFDirect$Motif_ID, rawPBM),TFDirect$Motif_ID),]

```

### download Protein data from Uniprot
```{r}
#uniprot can only handle certain number of searches at a time, so seperate all data into two parts
gene_ids <- unique(TFDirect[,c('TF_Name', 'TF_Species')])
searchUnitTemplate <- '(gene:"GENE" AND "SPECIES")'
searchQ <- gsub('GENE', gene_ids$TF_Name[1], searchUnitTemplate)
searchQ <- gsub('SPECIES', gene_ids$TF_Species[1], searchQ)
for(i in 1:nrow(gene_ids)){#nrow(gene_ids)
  addQ <- gsub('GENE', gene_ids$TF_Name[i], searchUnitTemplate)
  addQ <- gsub('SPECIES', gene_ids$TF_Species[i], addQ)
  searchQ <- paste0(searchQ, ' OR ', addQ)
}
```

### Import protein data 
```{r}
geneInfoUniprot <- read.delim(paste0(workDir,"geneInfoUniprot_all.tsv"))
geneInfoUniprot$Gene.Names <- paste0(geneInfoUniprot$Gene.Names, ' ')
```

### Match protein sequences data from uniprot with motif information from cisBP
```{r}

geneInfo <- data.frame()
for(i in 1:nrow(gene_ids)){
  add <- geneInfoUniprot[grep(paste0(gene_ids$TF_Name[i],' '), geneInfoUniprot$Gene.Names),]
  if(nrow(add) != 0){
    add <- add[nchar(add$Sequence) == max(nchar(add$Sequence)),][1,]
    addSeq <- add$Sequence
  }else{
    addSeq <- NA 
  }
  Motif <- unique(TFDirect[TFDirect$TF_Name == gene_ids$TF_Name[i],]$Motif_ID)
  addLine <- data.frame(name = gene_ids$TF_Name[i], seq = addSeq, motif = Motif)
  geneInfo <- rbind.data.frame(geneInfo, addLine)
}

geneInfo <- geneInfo[!is.na(geneInfo$seq),]
#write sequences as fasta to
writeFasta(geneInfo, paste0(workDir,'TF.fa'))
```

### Use hmmalign to align the output sequences with the Homeo Domain HMM model from pfam entry PF00010
HMMER 3.1b2 (February 2015); http://hmmer.org/
```{}
hmmalign -o cisBP_TF.sto HD.hmm TF.fa
```

### Use Jalview to hide non-hmm columns and output filtered residues, saved as TFhmm.sto
### Screen and Trim unusable alignment entries
```{r}
ali <- read.table(paste0(workDir,"TFhmm.sto"), quote="\"", fill = T)
ali <- ali[-nrow(ali),]
ali$V1 <- apply(ali, 1, function(x) strsplit(x[1],'/')[[1]][1])
ali$V1 <- apply(ali, 1, function(x) strsplit(x[1],'\\|')[[1]][2])
colnames(ali) <- c('name','alignment')
keep <- which(substr(ali$alignment, 50,50) == 'N')
cisBP_alignment <- ali[keep,]
geneInfo <- geneInfo[keep,]
#Check if cisBP_alignment and geneInfo matches
sum(cisBP_alignment$name != geneInfo$name) == 0
```

### Write Count Table for pyProBound motif modeling
```{r}
mkdirs(paste0(workDir,'CountTables'))
for(i in 1:nrow(geneInfo)){
  CT <- data.frame(name = Zscores$X8mer, score = Zscores[,geneInfo$motif[i]])
  rownames(CT) <- Zscores$X8mer
  colnames(CT) <- c('','0')
  CT <- CT[!is.na(CT$`0`),]
  CT <- CT[CT$`0` > 0,]
  fname <- geneInfo$motif[i]
  
  write.table(CT, file =paste0(workDir, 'CountTables/', fname, '_CT.tsv'), quote = F, dec = '.', row.names = F, col.names = T, sep = '\t')
}
```
### generate ProBound model with pyProBound
### Loading pwms from cisBP using geneInfo as index
```{r}
cisBP_motifs <- list()
options(warn=0)
for(i in 1:nrow(geneInfo)){
  motif <- geneInfo$motif[i]
  motif <- read.table(paste0(workDir, 'pwms/', motif, '.txt'), header = F)
  motif <- ddG2frequency(motif)
  motif[motif < 0.001] <- 0.001#add psuedo count
  rownames(motif) <- DNA()
  add <- list()
  add$gene_symbol <- geneInfo$name[i]
  add$study <- i
  add$mode <- 1
  add$matrix <- as.matrix(motif)
  cisBP_motifs[[length(cisBP_motifs)+1]] <- add
  add <- list()
  #add reverse compliment motif
  add$gene_symbol <- geneInfo$name[i]
  add$study <- i
  add$mode <- -1
  if(ncol(motif) == 0){
    rev <- motif
  }else{
    rev <- as.matrix(motif)[nrow(motif):1, ncol(motif):1]
  }
  rownames(rev) <- DNA()
  add$matrix <- rev
  cisBP_motifs[[length(cisBP_motifs)+1]] <- add
}
#check if length of cisBP_motif is equal to 2 times geneInfo
length(cisBP_motifs) == nrow(geneInfo)*2

```

### Load Homeo Domain 8mer seed of NNTRAYNN and score motifs with the seed
```{r}
#align to NNTDAYNN motif
HDseed <- matrix(nrow = 4,ncol = 8,data = c(0,0,0,0,
                                            0,0,0,0,
                                            0,0,0,1,
                                            1,0,1,1,
                                            1,0,0,0,
                                            0,1,0,1,
                                            0,0,0,0,
                                            0,0,0,0))
rownames(HDseed) <- DNA()

motifScores <- scoreMotifList(cisBP_motifs, HDseed, weight = c(0,0,1,1,1,1,0,0))
```

### Filter and reorder motif information according to motif matching score.
The motif (forward or reverse compliment) with the lowest motif matching score of each entry will be kept. For the same protein, entries are arranged from lower score to higher score
```{r}
idstds <- unique(motifScores[,1:2])
all_motifs <- data.frame(NULL)
for(i in 1:nrow(idstds)){
  fDT <- motifScores[motifScores$gene_symbol == idstds[i,1],]
  fDT <- fDT[fDT$study == idstds[i,2],]
  all_motifs <- rbind.data.frame(all_motifs, fDT[which.min(fDT$score),])
}
motifs_arranged <- data.frame(NULL)
motifScore <- c()
for(i in 1:length(unique(all_motifs$gene_symbol))){
  add <- all_motifs[all_motifs$gene_symbol == unique(all_motifs$gene_symbol)[i],]
  add <- dplyr::arrange(add, score)
  motifScore <- c(motifScore,add$score[1])
  motifs_arranged <- rbind.data.frame(motifs_arranged, add)
}
#Plot distribution of lowest motif matching score
plot.ecdf(as.numeric(motifScore))
```
```{r}
#Filter out motifs that do not match well with the seed
motifTable_baseline <- motifs_arranged[motifs_arranged$score < 1,]
#Number of unique HD protiens kept:
length(unique(motifTable_baseline$gene_symbol))
```

### Construct motif list and alignment table for Base Line (Motif similarity between the two best experiments)
```{r}
motifPoses <- c('N1','N2','T3','D4','A5','Y6','N7','N8')
cisBP_motifs_baseline <- filterMotifList(motifTable_baseline, cisBP_motifs, 8, motifPoses)
cisBP_alignment <- cisBP_alignment[match(motifTable_baseline$gene_symbol, cisBP_alignment$name),]
#check if motif List has same length as alignment
length(cisBP_motifs_baseline) == nrow(cisBP_alignment)
```

### Show base line accuracy
```{r}
evaMatrix <- matrix(nrow = 4, ncol = length(motifPoses), data = 0)
colnames(evaMatrix) <- motifPoses
for(pos in motifPoses){
  ScreenPoses <- c(pos,pos)
  BaseLines.ddG <- getBaseLineAccuracy(cisBP_motifs_baseline, pos = ScreenPoses, randomSample = F)
  colnames(BaseLines.ddG) <- c('pred','true')
  BaseLines.PFM <- predTrue.ddG2frequency(BaseLines.ddG, PFM = T)
  plotPredTrue(BaseLines.ddG, xlab = 'Experiment 1', ylab = 'Experiment 2', main = paste0('Base Line -ΔΔG/RT at ',pos))
  boxplot(groupedR2(BaseLines.ddG$pred, BaseLines.ddG$true, member = length(ScreenPoses)*4, mean= F), ylim = c(0,1),
        main = paste0(pos,' Mean R^2 = ', round(groupedR2(BaseLines.ddG$pred, BaseLines.ddG$true, 
                                                          member = length(ScreenPoses)*4, mean= T),4)))
  plotPredTrue(BaseLines.PFM, xlab = 'Experiment 1', ylab = 'Experiment 2', main = paste0('Base Line PFM at ',pos))
  hit <- sum(round(groupedR2(BaseLines.PFM$pred, BaseLines.PFM$true, member = 4, mean= F, throughZero = F)^(1/2),4) > 0.5)/(nrow(BaseLines.PFM)/4)
  boxplot(groupedR2(BaseLines.PFM$pred, BaseLines.PFM$true, member = length(ScreenPoses)*4, mean= F)^(1/2), ylim = c(0,1),
        main = paste0(pos, ' Mean PCC = ', round(groupedR2(BaseLines.PFM$pred, BaseLines.PFM$true, 
                                                     member = length(ScreenPoses)*4, mean= T, throughZero = F)^(1/2),4), ' Accuracy = ', round(hit,4)))
  evaMatrix[,pos] <- c(round(groupedR2(BaseLines.ddG$pred, 
        BaseLines.ddG$true, nrow(BaseLines.ddG), throughZero = T), 
        4), round(RMSD(BaseLines.ddG$pred, BaseLines.ddG$true), 
        4), round(groupedR2(BaseLines.PFM$pred, 
        BaseLines.PFM$true, nrow(BaseLines.PFM), throughZero = F), 
        4)^(1/2), round(hit,4))
}
evaMatrix_BL <- evaMatrix

```

### Make p-val table
```{r}
del <- which(duplicated(cisBP_alignment$alignment))
CV_alignment <- cisBP_alignment[-del,]
CV_motif <- cisBP_motifs_baseline[-del]
#genreate p-val tablle
pvalTable <- getPvalTable(CV_motif, Alignment = CV_alignment, pos_index = motifPoses)
#apply -log(10)
pvalTable <- -log(pvalTable,10)
#set NAs to 0
pvalTable[is.na(pvalTable)] <- 0
#plot heatmap
gplots::heatmap.2(as.matrix(as.data.frame(lapply(pvalTable, as.numeric))),dendrogram='none',
                  Rowv=FALSE, Colv=FALSE,trace='none',col = rev(heat.colors(12)), key.title = '-logP-val')
```

### Test on FamilyCode prediction with leave-one-out CV
Compare with paper: Wetzel, J. L., Zhang, K., & Singh, M. (2022). Learning probabilistic protein–DNA recognition codes from DNA-binding specificities using structural mappings. Genome Research, 32(9), 1776-1786.
Accuracy measurement in Figure 2
"PWM F, where entry fbj is the frequency with which nucleotide b is observed in column j in a set of aligned binding sites of that TF. "
```{r}
FCpredTrueList <- list()
for(pos in motifPoses){
  predTrue <- SVDregression.Iterative.CV(CV_motif,CV_alignment, pos = pos,Ftest_pVal = 0.001, useSimilarAA = T)
  #delete unsuccessful predictions
  predTrue <- predTrue[!is.na(predTrue$pred),]
  FCpredTrueList[[pos]] <- predTrue
  print(pos)
  
}
evaMatrix <- matrix(nrow = 4, ncol = length(motifPoses), data = 0)
colnames(evaMatrix) <- motifPoses
for(pos in motifPoses){
  predTrue <- FCpredTrueList[[pos]]
  #delete unsuccessful predictions
  predTrue <- predTrue[!is.na(predTrue$pred),]
  
  predTrue.PFM <- predTrue.ddG2frequency(predTrue, PFM = T)
  plotPredTrue(predTrue, main = paste0('FamilyCode prediction -ΔΔG/RT at ',pos))
  boxplot(groupedR2(predTrue$pred, predTrue$true, member = 4, mean= F), ylim = c(0,1),
        main = paste0(pos,' Mean R^2 = ', round(groupedR2(predTrue$pred, predTrue$true, 
                                                          member = 4, mean= T),4)))
  #plot confidance
  attr(predTrue, 'confidence') <- attr(predTrue, 'confidence')[attr(predTrue, 'confidence') <= 1]
  plot(attr(predTrue, 'confidence'),groupedR2(predTrue$pred, predTrue$true,member = 4, mean = F), ylab = 'Predict-True R^2', xlab = 'Confidence score', main = paste0('Confidence at ', pos))
  legend('bottomright', paste0('R^2 = ', round(groupedR2(groupedR2(predTrue$pred, predTrue$true,member = 4, mean = F),
                                                     attr(predTrue, 'confidence'), ),4)), bty = 'n')
  dt <- data.frame(conf = attr(predTrue, 'confidence'), R2 = groupedR2(predTrue$pred, predTrue$true,member = 4, mean = F))
  lm <- lm(R2~conf, data= dt)
  abline(lm, col= 'red')
  
  plotPredTrue(predTrue.PFM, main = paste0('FamilyCode prediction PFM at ',pos))
  hit <- sum(round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F, throughZero = F)^(1/2),4) > 0.5)/(nrow(predTrue.PFM)/4)
  boxplot(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F, throughZero = F)^(1/2), ylim = c(0,1),
        main = paste0(pos, ' Mean PCC = ', round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, 
                                                     member = 4, mean= T, throughZero = F)^(1/2),4), ' Accuracy = ', round(hit,4)))
  evaMatrix[,pos] <- c(round(groupedR2(predTrue$pred, 
        predTrue$true, nrow(predTrue), throughZero = T), 
        4), round(RMSD(predTrue$pred, predTrue$true), 
        4), round(groupedR2(predTrue.PFM$pred, 
        predTrue.PFM$true, nrow(predTrue.PFM), throughZero = F), 
        4)^(1/2), round(hit,4))
}
evaMatrix_FC <- evaMatrix
```
### Comparing prediction result with Closest sequence
from paper:
Lambert, S. A., Yang, A. W., Sasse, A., Cowley, G., Albu, M., Caddick, M. X., ... & Hughes, T. R. (2019). Similarity regression predicts evolution of transcription factor sequence specificity. Nature genetics, 51(6), 981-989.
Figure S6/S7
```{r}
evaMatrix <- matrix(nrow = 4, ncol = length(motifPoses), data = 0)
colnames(evaMatrix) <- motifPoses
for(pos in motifPoses){
  del <- which(duplicated(cisBP_alignment$alignment))
  CV_alignment <- cisBP_alignment[-del,]
  CV_motif <- cisBP_motifs_baseline[-del]
  predTrue <- closestSeqPred(CV_motif,CV_alignment, pos = pos)
  #delete unsuccessful predictions
  predTrue <- predTrue[!is.na(predTrue$pred),]
  predTrue.PFM <- predTrue.ddG2frequency(predTrue, PFM = T)
  
  plotPredTrue(predTrue, main = paste0('Closest sequence prediction -ΔΔG/RT at ',pos))
  boxplot(groupedR2(predTrue$pred, predTrue$true, member = 4, mean= F), ylim = c(0,1),
        main = paste0(pos,' Mean R^2 = ', round(groupedR2(predTrue$pred, predTrue$true, 
                                                          member = 4, mean= T),4)))
  
  plotPredTrue(predTrue.PFM, main = paste0('Closest sequence prediction PFM at ',pos))
  hit <- sum(round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F, throughZero = F)^(1/2),4) > 0.5)/(nrow(predTrue.PFM)/4)
  boxplot(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F, throughZero = F)^(1/2), ylim = c(0,1),
        main = paste0(pos, 'Mean PCC = ', round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, 
                                                     member = 4, mean= T, throughZero = F)^(1/2),4), ' Accuracy = ', round(hit,4)))
  
  evaMatrix[,pos] <- c(round(groupedR2(predTrue$pred, 
      predTrue$true, nrow(predTrue), throughZero = T), 
      4), round(RMSD(predTrue$pred, predTrue$true), 
      4), round(groupedR2(predTrue.PFM$pred, 
      predTrue.PFM$true, nrow(predTrue.PFM), throughZero = F), 
      4)^(1/2), round(hit,4))
}
evaMatrix_CS <- evaMatrix
```

### Comparing prediction result with Similarity regression
from paper:
Lambert, S. A., Yang, A. W., Sasse, A., Cowley, G., Albu, M., Caddick, M. X., ... & Hughes, T. R. (2019). Similarity regression predicts evolution of transcription factor sequence specificity. Nature genetics, 51(6), 981-989.
Figure S6/S7

```{r}
evaMatrix <- matrix(nrow = 4, ncol = length(motifPoses), data = 0)
colnames(evaMatrix) <- motifPoses
for(pos in motifPoses){
  del <- which(duplicated(cisBP_alignment$alignment))
  CV_alignment <- cisBP_alignment[-del,]
  CV_motif <- cisBP_motifs_baseline[-del]
  predTrue <- SRpred(CV_motif,CV_alignment, pos = pos, weightfile = 'cisBP_HD/F223_1.97d.json')
  #delete unsuccessful predictions
  predTrue <- predTrue[!is.na(predTrue$pred),]
  predTrue.PFM <- predTrue.ddG2frequency(predTrue, PFM = T)
  
  plotPredTrue(predTrue, main = paste0('Similarity regression prediction -ΔΔG/RT at ',pos))
  boxplot(groupedR2(predTrue$pred, predTrue$true, member = 4, mean= F), ylim = c(0,1),
        main = paste0(pos,' Mean R^2 = ', round(groupedR2(predTrue$pred, predTrue$true, 
                                                          member = 4, mean= T),4)))
  
  plotPredTrue(predTrue.PFM, main = paste0('Similarity regression prediction PFM at ',pos))
  hit <- sum(round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F, throughZero = F)^(1/2),4) > 0.5)/(nrow(predTrue.PFM)/4)
  boxplot(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F, throughZero = F)^(1/2), ylim = c(0,1),
        main = paste0(pos, 'Mean PCC = ', round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, 
                                                     member = 4, mean= T, throughZero = F)^(1/2),4), ' Accuracy = ', round(hit,4)))
  evaMatrix[,pos] <- c(round(groupedR2(predTrue$pred, 
    predTrue$true, nrow(predTrue), throughZero = T), 
    4), round(RMSD(predTrue$pred, predTrue$true), 
    4), round(groupedR2(predTrue.PFM$pred, 
    predTrue.PFM$true, nrow(predTrue.PFM), throughZero = F), 
    4)^(1/2), round(hit,4))
}
evaMatrix_SR <- evaMatrix

```
```{r}
evaMatrix <- rbind.data.frame(evaMatrix_BL, evaMatrix_FC,  evaMatrix_SR, evaMatrix_CS)
evaMatrix <- evaMatrix[c((1:4)*4-3, (1:4)*4-2, (1:4)*4-1,(1:4)*4), ]
```

