---
title: "FamilyCode on CisBP data (HD proteins as example)"
output: html_notebook
---
## Family Code application on combinatory data in CisBP. Using HD proteins as an example
### Import library and data
```{r}
library(FamilyCode)
workDir <- 'cisBP_HD/'
#download from cisBP 'bulk download' tab, incuding pwms and TF_Information
TF_Information <- read.delim("cisBP_HD/TF_Information.txt")
TFDirect <- TF_Information[TF_Information$TF_Status == 'D',]

#import downloaded data from uniprot
#geneInfoUniprot <- read.delim("cisBP_HD/uniprotkb_Homeo_Domain_2024_05_22.tsv")
#geneInfoUniprot$Gene.Names <- paste0(geneInfoUniprot$Gene.Names, ' ')
```

### download Protein data from Uniprot
```{r}
#download from unprot, searching for 'Homeo Domain'
#uniprot can only handle certain number of searches at a time, so seperate all data into two parts
gene_ids <- unique(TFDirect[,c('TF_Name', 'TF_Species')])
searchUnitTemplate <- '(gene:"GENE" AND "SPECIES")'
searchQ <- gsub('GENE', gene_ids$TF_Name[1], searchUnitTemplate)
searchQ <- gsub('SPECIES', gene_ids$TF_Species[1], searchQ)
for(i in 1:330){#nrow(gene_ids)
  addQ <- gsub('GENE', gene_ids$TF_Name[i], searchUnitTemplate)
  addQ <- gsub('SPECIES', gene_ids$TF_Species[i], addQ)
  searchQ <- paste0(searchQ, ' OR ', addQ)
}
cat(searchQ)
```

```{r}
searchQ <- gsub('GENE', gene_ids$TF_Name[1], searchUnitTemplate)
searchQ <- gsub('SPECIES', gene_ids$TF_Species[1], searchQ)
for(i in 331:nrow(gene_ids)){#nrow(gene_ids)
  addQ <- gsub('GENE', gene_ids$TF_Name[i], searchUnitTemplate)
  addQ <- gsub('SPECIES', gene_ids$TF_Species[i], addQ)
  searchQ <- paste0(searchQ, ' OR ', addQ)
}
cat(searchQ)
```

### Import protein data 
```{r}
geneInfoUniprot1 <- read.delim("cisBP_HD/geneInfoUniprot_all1.tsv")
geneInfoUniprot2 <- read.delim("cisBP_HD/geneInfoUniprot_all2.tsv")
geneInfoUniprot <- rbind.data.frame(geneInfoUniprot1, geneInfoUniprot2)
geneInfoUniprot$Gene.Names <- paste0(geneInfoUniprot$Gene.Names, ' ')
```


### Match protein sequences data from uniprot with motif information from cisBP
```{r}

geneInfo <- data.frame()
for(i in 1:nrow(gene_ids)){
  add <- geneInfoUniprot[grep(paste0(gene_ids$TF_Name[i],' '), geneInfoUniprot$Gene.Names),]
  if(nrow(add) != 0){
    add <- add[nchar(add$Sequence) == max(nchar(add$Sequence)),][1,]
    addSeq <- add$Sequence
  }else{
    addSeq <- NA 
  }
  Motif <- unique(TFDirect[TFDirect$TF_Name == gene_ids$TF_Name[i],]$Motif_ID)
  addLine <- data.frame(name = gene_ids$TF_Name[i], seq = addSeq, motif = Motif)
  geneInfo <- rbind.data.frame(geneInfo, addLine)
}

geneInfo <- geneInfo[!is.na(geneInfo$seq),]
#write sequences as fasta to
writeFasta(geneInfo, 'cisBP_HD/TF.fa')
```
### Use hmmalign to align the output sequences with the Homeo Domain HMM model from pfam entry PF00046
```{}
hmmalign -o cisBP_TF.sto HD.hmm TF.fa
```

### Use Jalview to hide non-hmm columns and output filtered residues, saved as TFhmm.sto
### Screen and Trim unusable alignment entries
```{r}
ali <- read.table(r"(cisBP_HD/TFhmm.sto)", quote="\"", fill = T)
ali <- ali[-nrow(ali),]
ali$V1 <- apply(ali, 1, function(x) strsplit(x[1],'/')[[1]][1])
ali$V1 <- apply(ali, 1, function(x) strsplit(x[1],'\\|')[[1]][2])
colnames(ali) <- c('name','alignment')
keep <- which(substr(ali$alignment, 50,50) == 'N')
cisBP_alignment <- ali[keep,]
geneInfo <- geneInfo[keep,]
#Check if cisBP_alignment and geneInfo matches
sum(cisBP_alignment$name != geneInfo$name) == 0
```

### Loading pwms from cisBP using geneInfo as index
```{r}
cisBP_motifs <- list()
for(i in 1:nrow(geneInfo)){
  motif <- geneInfo$motif[i]
  motif <- t(read.table(paste0('cisBP_HD/pwms/', motif, '.txt'), header = T, row.names = 1))
  motif[motif < 0.001] <- 0.001#add psuedo count
  add <- list()
  add$gene_symbol <- geneInfo$name[i]
  add$study <- i
  add$mode <- 1
  add$matrix <- as.matrix(motif)
  cisBP_motifs[[length(cisBP_motifs)+1]] <- add
  add <- list()
  #add reverse compliment motif
  add$gene_symbol <- geneInfo$name[i]
  add$study <- i
  add$mode <- -1
  if(ncol(motif) == 0){
    rev <- motif
  }else{
    rev <- as.matrix(motif)[nrow(motif):1, ncol(motif):1]
  }
  rownames(rev) <- DNA()
  add$matrix <- rev
  cisBP_motifs[[length(cisBP_motifs)+1]] <- add
}
#check if length of cisBP_motif is equal to 2 times geneInfo
length(cisBP_motifs) == nrow(geneInfo)*2
```

### Load Homeo Domain 8mer seed of NNTRAYNN and score motifs with the seed
```{r}
#align to NNTRAYNN motif, would need to change if working with other motif
HDseed <- matrix(nrow = 4,ncol = 8,data = c(0,0,0,0,
                                            0,0,0,0,
                                            0,0,0,1,
                                            1,0,1,1,
                                            1,0,0,0,
                                            0,1,0,1,
                                            0,0,0,0,
                                            0,0,0,0))
rownames(HDseed) <- DNA()

#Scoring would result in some motif shorter than the seed, those motifs will be discarded. 
motifScores <- scoreMotifList(cisBP_motifs, HDseed, weight = c(0,0,1,1,3,1,0,0))
```

### Filter and reorder motif information according to motif matching score.
The motif (forward or reverse compliment) with the lowest motif matching score of each entry will be kept. For the same protein, entries are arranged from lower score to higher score
```{r}
idstds <- unique(motifScores[,1:2])
all_motifs <- data.frame(NULL)
for(i in 1:nrow(idstds)){
  fDT <- motifScores[motifScores$gene_symbol == idstds[i,1],]
  fDT <- fDT[fDT$study == idstds[i,2],]
  all_motifs <- rbind.data.frame(all_motifs, fDT[which.min(fDT$score),])
}
motifs_arranged <- data.frame(NULL)
motifScore <- c()
for(i in 1:length(unique(all_motifs$gene_symbol))){
  add <- all_motifs[all_motifs$gene_symbol == unique(all_motifs$gene_symbol)[i],]
  add <- dplyr::arrange(add, score)
  motifScore <- c(motifScore,add$score[1])
  motifs_arranged <- rbind.data.frame(motifs_arranged, add)
}
#Plot distribution of lowest motif matching score
plot.ecdf(as.numeric(motifScore))
```

```{r}
#Filter out motifs that do not match well with the seed
motifTable_baseline <- motifs_arranged[motifs_arranged$score < 0.5,]
#Number of unique HD protiens kept:
length(unique(motifTable_baseline$gene_symbol))
```
### Construct motif list and alignment table for Base Line (Motif similarity between the two best experiments)
```{r}
motifPoses <- c('N1','N2','T3','D4','A5','Y6','N7','N8')
cisBP_motifs_baseline <- filterMotifList(motifTable_baseline, cisBP_motifs, 8, motifPoses)
cisBP_alignment <- cisBP_alignment[match(motifTable_baseline$gene_symbol, cisBP_alignment$name),]
#check if motif List has same length as alignment
length(cisBP_motifs_baseline) == nrow(cisBP_alignment)
del <- which(duplicated(cisBP_alignment$name))
CV_alignment <- cisBP_alignment[-del,]
CV_motifs <- cisBP_motifs_baseline[-del]
```
### Show base line accuracy
```{r}
for(pos in motifPoses){
  ScreenPoses <- c(pos,pos)
  BaseLines.ddG <- getBaseLineAccuracy(cisBP_motifs_baseline, pos = ScreenPoses, randomSample = F)
  colnames(BaseLines.ddG) <- c('pred','true')
  BaseLines.PFM <- predTrue.ddG2frequency(BaseLines.ddG, PFM = T)
  plotPredTrue(BaseLines.ddG, xlab = 'Experiment 1', ylab = 'Experiment 2', main = paste0('Base Line -ddG/RT at ',pos))
  boxplot(groupedR2(BaseLines.ddG$pred, BaseLines.ddG$true, member = length(ScreenPoses)*4, mean= F), ylim = c(0,1),
        main = paste0(pos,' Mean R^2 = ', round(groupedR2(BaseLines.ddG$pred, BaseLines.ddG$true, 
                                                          member = length(ScreenPoses)*4, mean= T),4)))
  plotPredTrue(BaseLines.PFM, xlab = 'Experiment 1', ylab = 'Experiment 2', main = paste0('Base Line PFM at ',pos))
  hit <- sum(round(groupedR2(BaseLines.PFM$pred, BaseLines.PFM$true, member = 4, mean= F)^(1/2),4) > 0.5)/(nrow(BaseLines.PFM)/4)
  boxplot(groupedR2(BaseLines.PFM$pred, BaseLines.PFM$true, member = length(ScreenPoses)*4, mean= F)^(1/2), ylim = c(0,1),
        main = paste0(pos, ' Mean PCC = ', round(groupedR2(BaseLines.PFM$pred, BaseLines.PFM$true, 
                                                     member = length(ScreenPoses)*4, mean= T)^(1/2),4), ' Accuracy = ', round(hit,4)))
  
}

```
### Make p-val table
```{r}
#genreate p-val tablle
pvalTable <- getPvalTable(CV_motifs, Alignment = CV_alignment, pos_index = motifPositions)
#apply -log(10)
pvalTable <- -log(pvalTable,10)
#set NAs to 0
pvalTable[is.na(pvalTable)] <- 0
#plot heatmap
gplots::heatmap.2(as.matrix(as.data.frame(lapply(pvalTable, as.numeric))),dendrogram='none',
                  Rowv=FALSE, Colv=FALSE,trace='none',col = rev(heat.colors(12)), key.title = '-logP-val')
```

### Test on FamilyCode prediction with leave-one-out CV
Compare with paper: Wetzel, J. L., Zhang, K., & Singh, M. (2022). Learning probabilistic protein–DNA recognition codes from DNA-binding specificity using structural mappings. Genome Research, 32(9), 1776-1786.
Accuracy measurement in Figure 2
"PWM F, where entry fbj is the frequency with which nucleotide b is observed in column j in a set of aligned binding sites of that TF. "
```{r}
FCpredTrueList <- list()
for(pos in motifPoses){
  predTrue <- SVDregression.Iterative.CV(CV_motifs,CV_alignment, pos = pos,Ftest_pVal = 0.001, useSimilarAA = T)
  #delete unsuccessful predictions
  predTrue <- predTrue[!is.na(predTrue$pred),]
  FCpredTrueList[[pos]] <- predTrue
  print(pos)
  
}

#plot results
mkdirs(paste0(workDir, 'plots'))
for(pos in motifPoses){
  predTrue <- FCpredTrueList[[pos]]
  #delete unsuccessful predictions
  #predTrue <- predTrue[!is.na(predTrue$pred),]
  attr(predTrue, 'confidence') <- attr(predTrue, 'confidence')[attr(predTrue, 'confidence') <= 1]
  predTrue.PFM <- predTrue.ddG2frequency(predTrue, PFM = T)
  plotPredTrue(predTrue, main = paste0('FamilyCode prediction -ddG/RT at ',pos))
  boxplot(groupedR2(predTrue$pred, predTrue$true, member = 4, mean= F), ylim = c(0,1),main = paste0(pos,' Mean R^2 = ', round(groupedR2(predTrue$pred, predTrue$true, member = 4, mean= T),4)))
  #plot confidance
  plot(attr(predTrue, 'confidence'),groupedR2(predTrue$pred, predTrue$true,member = 4, mean = F), ylab = 'Predict-True R^2', xlab = 'Confidence score', main = paste0('Confidence at ', pos))
  legend('bottomright', paste0('R^2 = ', round(groupedR2(groupedR2(predTrue$pred, predTrue$true,member = 4, mean = F),
                                                     attr(predTrue, 'confidence'), ),4)), bty = 'n')
  dt <- data.frame(conf = attr(predTrue, 'confidence'), R2 = groupedR2(predTrue$pred, predTrue$true,member = 4, mean = F))
  lm <- lm(R2~conf, data= dt)
  abline(lm, col= 'red')
  #plot frequency

  plotPredTrue(predTrue.PFM, main = paste0('FamilyCode prediction PFM at ',pos))
  hit <- sum(round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F)^(1/2),4) > 0.5)/(nrow(predTrue.PFM)/4)
  boxplot(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F)^(1/2), ylim = c(0,1),main = paste0(pos, ' Mean PCC = ', round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= T)^(1/2),4), ' Accuracy = ', round(hit,4)))
}
```

### Comparing prediction result with Closest sequence 
from paper:
Weirauch, M. T., Yang, A., Albu, M., Cote, A. G., Montenegro-Montero, A., Drewe, P., ... & Hughes, T. R. (2014). Determination and inference of eukaryotic transcription factor sequence specificity. Cell, 158(6), 1431-1443.
```{r}
for(pos in motifPoses){
  del <- which(duplicated(cisBP_alignment$alignment))
  CV_alignment <- cisBP_alignment[-del,]
  CV_motif <- cisBP_motifs_baseline[-del]
  predTrue <- closestSeqPred(CV_motif,CV_alignment, pos = pos)
  #delete unsuccessful predictions
  predTrue <- predTrue[!is.na(predTrue$pred),]
  predTrue.PFM <- predTrue.ddG2frequency(predTrue, PFM = T)
  
  plot.ecdf(predTrue$similarity)
  plotPredTrue(predTrue, main = paste0('Closest Sequence prediction -ddG/RT at ',pos))
  boxplot(groupedR2(predTrue$pred, predTrue$true, member = 4, mean= F), ylim = c(0,1),
        main = paste0(pos,' Mean R^2 = ', round(groupedR2(predTrue$pred, predTrue$true, 
                                                          member = 4, mean= T),4)))
  
  plotPredTrue(predTrue.PFM, main = paste0('Closest Sequence prediction PFM at ',pos))
  hit <- sum(round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F)^(1/2),4) > 0.5)/(nrow(predTrue.PFM)/4)
  boxplot(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F)^(1/2), ylim = c(0,1),
        main = paste0(pos, 'Mean PCC = ', round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, 
                                                     member = 4, mean= T)^(1/2),4), ' Accuracy = ', round(hit,4)))
  
}
```

### Comparing prediction result with Similarity regression
from paper:
Lambert, S. A., Yang, A. W., Sasse, A., Cowley, G., Albu, M., Caddick, M. X., ... & Hughes, T. R. (2019). Similarity regression predicts evolution of transcription factor sequence specificity. Nature genetics, 51(6), 981-989.
Figure S6/S7
```{r}
for(pos in motifPoses){
  del <- which(duplicated(cisBP_alignment$alignment))
  CV_alignment <- cisBP_alignment[-del,]
  CV_motif <- cisBP_motifs_baseline[-del]
  predTrue <- SRpred(CV_motif,CV_alignment, pos = pos, weightfile = 'cisBP_HD/F223_1.97d.json')
  #delete unsuccessful predictions
  predTrue <- predTrue[!is.na(predTrue$pred),]
  predTrue.PFM <- predTrue.ddG2frequency(predTrue, PFM = T)
  
  plot.ecdf(predTrue$similarity)
  plotPredTrue(predTrue, main = paste0('Similarity regression prediction -ddG/RT at ',pos))
  boxplot(groupedR2(predTrue$pred, predTrue$true, member = 4, mean= F), ylim = c(0,1),
        main = paste0(pos,' Mean R^2 = ', round(groupedR2(predTrue$pred, predTrue$true, 
                                                          member = 4, mean= T),4)))
  
  plotPredTrue(predTrue.PFM, main = paste0('Similarity regression prediction PFM at ',pos))
  hit <- sum(round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F)^(1/2),4) > 0.5)/(nrow(predTrue.PFM)/4)
  boxplot(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F)^(1/2), ylim = c(0,1),
        main = paste0(pos, 'Mean PCC = ', round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, 
                                                     member = 4, mean= T)^(1/2),4), ' Accuracy = ', round(hit,4)))
}

```
### Similarity regression of all 8 posiitons combined
```{r}
pos <- motifPoses
del <- which(duplicated(cisBP_alignment$alignment))
CV_alignment <- cisBP_alignment[-del,]
CV_motif <- cisBP_motifs_baseline[-del]
predTrue <- SRpred(CV_motif,CV_alignment, pos = pos, weightfile = 'cisBP_HD/F223_1.97d.json')
#delete unsuccessful predictions
predTrue <- predTrue[!is.na(predTrue$pred),]
predTrue.PFM <- predTrue.ddG2frequency(predTrue, PFM = T)

plot.ecdf(predTrue$similarity)
plotPredTrue(predTrue, main = paste0('Similarity regression prediction -ddG/RT at ',pos))
boxplot(groupedR2(predTrue$pred, predTrue$true, member = 4, mean= F), ylim = c(0,1),
      main = paste0(pos,' Mean R^2 = ', round(groupedR2(predTrue$pred, predTrue$true, 
                                                        member = 4, mean= T),4)))

plotPredTrue(predTrue.PFM, main = paste0('Similarity regression prediction PFM at ',pos))
hit <- sum(round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F)^(1/2),4) > 0.5)/(nrow(predTrue.PFM)/4)
boxplot(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, member = 4, mean= F)^(1/2), ylim = c(0,1),
      main = paste0(pos, 'Mean PCC = ', round(groupedR2(predTrue.PFM$pred, predTrue.PFM$true, 
                                                   member = 4, mean= T)^(1/2),4), ' Accuracy = ', round(hit,4)))
```

### systemetic test of prediction methods with subsampling
```{r}
options(warn=0)
CV_alignment <- cisBP_alignment[-del,]
CV_motif <- cisBP_motifs_baseline[-del]
nset <- nrow(CV_alignment)
results <- matrix(nrow = 6, ncol = 10, data = 0)
for(i in 2:10){
  FCR2List <- c()
  FCrmsdList <- c()
  CSR2List <- c()
  CSrmsdList <- c()
  SRR2List <- c()
  SRrmsdList <- c()
  for(j in 1:20){
    trainID <- sample(1:nset, floor(nset-nset/(i)))
    train_alignment <- CV_alignment[trainID,]
    train_motifs <- CV_motif[trainID]
    test_alignment <- CV_alignment[-trainID,]
    test_motifs <- CV_motif[-trainID]
    #FC
    svdModel <- makeSVDModel(train_motifs, train_alignment, Positions = c('N1','N2','T3','R4','A5','Y6','N7','N8'))
    pred <- lapply(test_alignment$alignment, function(x) SVDmodelPredPSAM(svdModel, x))
    pred <- lapply(pred, frequency2ddG)
    true <- lapply(test_motifs, function(x) x$matrix)
    true <- lapply(true, frequency2ddG)
    predTrue <- data.frame(pred = as.numeric(unlist(pred)), true = as.numeric(unlist(true)))
    FCR2 <- round(groupedR2(predTrue$true, predTrue$pred),4)
    FCrmsd <- round(RMSD(predTrue$true,predTrue$pred),4)
    #CS
    pred_motifs <- CSpredPSAM(test_alignment, train_motifs, train_alignment)
    pred <- lapply(pred_motifs, function(x) x$matrix)
    pred <- lapply(pred, frequency2ddG)
    true <- lapply(test_motifs, function(x) x$matrix)
    true <- lapply(true, frequency2ddG)
    predTrue <- data.frame(pred = as.numeric(unlist(pred)), true = as.numeric(unlist(true)))
    CSR2 <- round(groupedR2(predTrue$true, predTrue$pred),4)
    CSrmsd <- round(RMSD(predTrue$true,predTrue$pred),4)
    #SR
    pred_motifs <- SRpredPSAM(test_alignment, train_motifs, train_alignment, weightfile = 'cisBP_HD/F223_1.97d.json')
    pred <- lapply(pred_motifs, function(x) x$matrix)
    pred <- lapply(pred, frequency2ddG)
    true <- lapply(test_motifs, function(x) x$matrix)
    true <- lapply(true, frequency2ddG)
    predTrue <- data.frame(pred = as.numeric(unlist(pred)), true = as.numeric(unlist(true)))
    SRR2 <- round(groupedR2(predTrue$true, predTrue$pred),4)
    SRrmsd <- round(RMSD(predTrue$true,predTrue$pred),4)
    FCR2List <- c(FCR2List, FCR2)
    FCrmsdList <- c(FCrmsdList, FCrmsd)
    CSR2List <- c(CSR2List, CSR2)
    CSrmsdList <- c(CSrmsdList, CSrmsd)
    SRR2List <- c(SRR2List, SRR2)
    SRrmsdList <- c(SRrmsdList, SRrmsd)
  }
  results[1,i] <- mean(FCR2List)
  results[2,i] <- mean(FCrmsdList)
  results[3,i] <- mean(CSR2List)
  results[4,i] <- mean(CSrmsdList)
  results[5,i] <- mean(SRR2List)
  results[6,i] <- mean(SRrmsdList)
}
results <- results[,2:10]
matplot(1-1/2:10, t(results[c(1,3,5),]), type = "l", lty = 1, 
        col = c("red", "blue", "green"), xlab = "X", 
        ylab = "Y", main = "Multiple Lines Plot")
matplot(1-1/2:10, t(results[c(2,4,6),]), type = "l", lty = 1, 
        col = c("red", "blue", "green"), xlab = "X", 
        ylab = "Y", main = "Multiple Lines Plot")
```

